### 2017-12-31 Analysis/RSV.Rmd
Added RSV analysis code

---
title: "RSV"
output: html_notebook
---

```{r setup, include=FALSE, error=FALSE, warning=FALSE, message=FALSE}
install.packages(c("dplyr", "zipcode", "reshape", "mgcv", "ggplot2", "gridExtra", "devtools"))
library(devtools)
install_github("gavinsimpson/tsgam")
install_github("kassambara/easyGgplot2")
install_github("airbornemint/outbreak-predict")

library(reshape)
library(plyr)
library(dplyr)
library(zipcode)
library(mgcv)
library(tsgam)
library(ggplot2)
library(easyGgplot2)
library(utils)
library(outbreakpredict)
data(zipcode)
set.seed(0)
```

```{r}
rsvWeek = function(weeki) {
  (weeki + 26) %% 52 - 26
}

dataset = read.csv("../ct rsv time series.csv", header=TRUE, colClasses=c("patzip"="character"))
# Add week-of-year, with zero = end of June, drop first 5.5 years because of coding differences
dataset = dataset %>%
  mutate(weekiadj=round(dataset$weeki * 364 / 365.25), weekofyear=rsvWeek(weekiadj), year=1990 + (weekiadj + 26) %/% 52) %>%
  filter(weeki > 5.5*52) %>%
  filter(weeki < 22.5*52)
# Add state based on zipcode and drop everything not in CT
zipState = zipcode %>% select(zip, state)
dataset = dataset %>% 
  left_join(zipcode %>% select(zip, state), by=c("patzip"="zip")) %>% 
  filter(state=="CT")
# Add county based on zipcode
countyByZip = read.csv("../ct counties.csv", header=TRUE, colClasses=c("zip"="character"))
counties = countyByZip %>% select(county) %>% distinct()
dataset = dataset %>% inner_join(countyByZip, by=c("patzip"="zip"))

# Sort counties by total RSV
totalRSVByCounty = dataset %>%
  group_by(county) %>% 
  summarize(total_rsv=sum(rsv)) %>%
  arrange(total_rsv) %>%
  mutate(county=factor(county, county))

rsvCounties = totalRSVByCounty$county
rsvYears = sort(unique(dataset$year))
rsvWeekRange = sort(unique(dataset$weeki))
rsvWeekOfYearRange = sort(unique(dataset$weekofyear))

dataset = dataset %>%
  mutate(county=factor(county, levels=levels(rsvCounties)))

```

```{r}
ggplot(dataset %>% 
         group_by(weeki) %>%
         summarize(rsv=sum(rsv)), 
       aes(x=weeki, y=rsv)) +
  geom_hline(aes(yintercept = 0)) +
  geom_line() +
  labs(y="RSV cases/week", x="Time", title="All counties")

plots = lapply(rsvCounties, function(c) {
  ggplot(dataset %>% 
           filter(county==c) %>%
           group_by(weeki) %>%
           summarize(rsv=sum(rsv)), 
         aes(x=weeki, y=rsv)) +
    geom_hline(aes(yintercept = 0)) +
    geom_line() +
    labs(y="RSV cases/week", x="Time", title=c)
})

ggplot2.multiplot(plotlist=plots, cols=4)

```

```{r include=FALSE}
# Reshape counties into columns and add column for total
rsvByCountyAndWeekOfYear = dataset %>%
  group_by(county, weekofyear) %>%
  summarize(rsv=sum(rsv)) %>%
  group_by(weekofyear) %>%
  do(rbind(., as.data.frame(.) %>%
             summarize(county="total", weekofyear=max(weekofyear), rsv=sum(rsv))
           ))

rsvForAllCountiesByWeekOfYear = rsvByCountyAndWeekOfYear %>%
  melt(id=c("weekofyear", "county"), measured=c("rsv")) %>%
  cast(weekofyear ~ county + variable) %>%
  rename_all(funs(gsub("_rsv", "", .)))

maxRSVByCounty = rsvByCountyAndWeekOfYear %>%
  group_by(county) %>% 
  summarize(max_rsv=max(rsv))

rsvRelForAllCountiesByWeekOfYear = rsvByCountyAndWeekOfYear %>%
  inner_join(maxRSVByCounty, by="county") %>%
  mutate(rsv_rel=rsv/max_rsv) %>%
  select(-max_rsv, -rsv) %>%
  melt(id=c("weekofyear", "county"), measured=c("rsv_rel")) %>%
  cast(weekofyear ~ county + variable)

allCountiesTS = rsvForAllCountiesByWeekOfYear %>%
  arrange(weekofyear) %>%
  select(-weekofyear) %>%
  ts(frequency=1, start=c(-26, 1))

allCountiesRelTS = rsvRelForAllCountiesByWeekOfYear %>%
  arrange(weekofyear) %>%
  select(-weekofyear) %>%
  ts(frequency=1, start=c(-26, 1))

plot.ts(allCountiesTS, nc=1)
plot.ts(allCountiesRelTS, nc=1, plot.type="single", col=rainbow(8), lwd=2)
```

```{r include=FALSE}
# Drop counties in which max # of RSV cases in a week is <100
maxRSVByInterestingCounty = maxRSVByCounty %>% filter(max_rsv > 50)

rsvForInterestingCountiesByWeekOfYear = rsvByCountyAndWeekOfYear %>%
  inner_join(maxRSVByInterestingCounty, by="county") %>%
  select(-max_rsv) %>%
  melt(id=c("weekofyear", "county"), measured=c("rsv", "rsv_rel")) %>%
  cast(weekofyear ~ county + variable)

rsvRelForInterestingCountiesByWeekOfYear = rsvByCountyAndWeekOfYear %>%
  inner_join(maxRSVByInterestingCounty, by="county") %>%
  mutate(rsv_rel=rsv/max_rsv) %>%
  select(-max_rsv, -rsv) %>%
  melt(id=c("weekofyear", "county"), measured=c("rsv_rel")) %>%
  cast(weekofyear ~ county + variable)

interestingCountiesTS = rsvForInterestingCountiesByWeekOfYear %>%
  arrange(weekofyear) %>%
  select(-weekofyear) %>%
  ts(frequency=1, start=c(-26, 1))

interestingCountiesRelTS = rsvRelForInterestingCountiesByWeekOfYear %>%
  arrange(weekofyear) %>%
  select(-weekofyear) %>%
  ts(frequency=1, start=c(-26, 1))

plot.ts(interestingCountiesTS, nc=1)
plot.ts(interestingCountiesRelTS, nc=1, plot.type="single", col=rainbow(5), lwd=2)
```

```{r}
outbreak.calc.unprotected = function(model, params, time, strategies) {
  # Get model predictions for given (randomized) param values
  time = time + 0.5
  predictors = model %>% predict(data.frame(time=time), type="lpmatrix")
  fit = predictors %*% params

  # Map spline fit back to data
  fit = fit %>% model$family$linkinv()

  # Calculate total # of unprotected cases for each strategy
  unprotected = as.data.frame(lapply(strategies, function(strat) {
    sum(fit * (1 - sapply(time, strat)))
  }))

  # Calculate total # of cases 
  total = sum(fit)

  unprotected %>% 
    rename_all(funs(
      sprintf("%s.count", .)
    )) %>%
    cbind(
      (unprotected / total) %>%
        rename_all(funs(
          sprintf("%s.frac", .)
        ))
    ) %>% 
    mutate(total=total)
}
```

```{r}
eps = .05
seasonThreshold = 0.025
rsvTime = dataset$weekofyear
modelTime = data.frame(time=seq(min(rsvTime) - 1 + eps, max(rsvTime), eps))

rsvCountyGroups = data.frame(county=rsvCounties) %>%
  mutate(
    group.all=TRUE,
    group.lowIncidence=county %in% c("Tolland", "Windham", "Middlesex", "Litchfield")
  )

ppxDuration = 24 # weeks

evalStrategy = function(start, end, time) {
  as.numeric((time > rsvWeek(start)) && (time < rsvWeek(end)))
}

evalOnsetStrategy = function(start, time) {
  as.numeric((time > rsvWeek(start)) && (time < rsvWeek(start + ppxDuration)))
}

evalOffsetStrategy = function(end, time) {
  as.numeric((time > rsvWeek(end - ppxDuration)) && (time < rsvWeek(end)))
}

ppxFixedStrategies = function(c=NULL) {
  stateOnset = stateThresholds$onset.median
  stateOffset = stateThresholds$offset.median

  if (!is.null(c)) {
    countyOnset = (thresholdsByCounty %>% filter(county==c))$onset.median
    countyOffset = (thresholdsByCounty %>% filter(county==c))$offset.median
  }

  list(
    # never=function(time) { 0 },
    # always=function(time) { 1 },
    aap=function(time) {
      evalOnsetStrategy(strptime("11-15", format="%m-%d")$yday / 7, time)
    },
    stateOnset=function(time) {
      evalOnsetStrategy(stateOnset, time)
    },
    stateOffset=function(time) {
      evalOffsetStrategy(stateOffset, time)
    },
    countyOnset=function(time) {
      evalOnsetStrategy(countyOnset, time)
    },
    countyOffset=function(time) {
      evalOffsetStrategy(countyOffset, time)
    }
  )
}

ppxSlidingStrategies = function(c=NULL, y=NULL) {
  if (!is.null(y) && !is.null(c)) {
    stateSlidingOnset = (stateThresholdsByWindow %>% filter(year==y))$onset.median
    stateSlidingOffset = (stateThresholdsByWindow %>% filter(year==y))$offset.median
  }

  list(
    stateSlidingOnset=function(time) {
      evalOnsetStrategy(stateSlidingOnset, time)
    },
    stateSlidingOffset=function(time) {
      evalOffsetStrategy(stateSlidingOffset, time)
    }
  )
}

slidingWindowDuration = 3 # Years

```

```{r}
stateObs = dataset %>% 
  rename(time=weekofyear) %>% 
  group_by(time) %>% 
  summarize(rsv=sum(rsv)) %>% 
  mutate(rsv.cum=cumsum(rsv), rsv.cum.frac=rsv.cum / sum(rsv)) %>% 
  as.data.frame()

stateModel = gam(rsv ~ s(time, k=20, bs="cp", m=3), family=poisson, data=stateObs)

statePred = stateModel %>% 
  predict(type="response", newdata=modelTime, se.fit=TRUE)

statePred = modelTime %>% 
  cbind(data.frame(rsv.fit=statePred$fit, rsv.fit.se=statePred$se.fit)) %>%
  cbind(stateModel %>% outbreak.predict.timeseries(modelTime, function(model, params, newdata) { outbreak.calc.cum(model, params, newdata) })) %>%
  rename(rsv.cum.fit.lower=lower, rsv.cum.fit.upper=upper, rsv.cum.fit=median)

stateThresholds = stateModel %>%
  outbreak.predict.scalars(
    modelTime,
    function(model, params, newdata) { 
      outbreak.calc.thresholds(model, params, newdata$time, onset=seasonThreshold, offset=1 - seasonThreshold)
    }
  )

summarizeCountyGroups = function(data) {
  rbind(
      as.data.frame(data) %>%
        select(county, time, rsv), 
      as.data.frame(data) %>%
        filter(group.all==TRUE) %>%
        summarize(county="all", time=max(time), rsv=sum(rsv)),
      as.data.frame(data) %>%
        filter(group.lowIncidence==TRUE) %>%
        summarize(county="lowIncidence", time=max(time), rsv=sum(rsv))
    )  
}
```

```{r}
stateObsByWindow = dataset %>% 
  mutate(time=weeki) %>% 
  group_by(time, year, weekofyear) %>% 
  summarize(rsv=sum(rsv)) %>%
  as.data.frame()

total = data.frame(time=stateObsByWindow$time, rsv=0)

for (idx in seq(1, slidingWindowDuration)) {
  total = total %>%
    inner_join(stateObsByWindow %>%
                 mutate(time=time + 52 * idx) %>%
                 select(time, rsv),
               by="time"
               ) %>%
    mutate(rsv=rsv.x + rsv.y) %>%
    select(time, rsv)
}

stateObsByWindow = stateObsByWindow %>%
  select(-rsv) %>%
  inner_join(total, by="time") %>%
  group_by(year) %>%
  mutate(rsv.cum=cumsum(rsv), rsv.cum.frac=rsv.cum / sum(rsv)) %>% 
  select(-time) %>%
  rename(time=weekofyear) %>%
  filter(year >= min(rsvYears) + slidingWindowDuration) %>%
  as.data.frame()

rsvWindowYears = unique(stateObsByWindow$year)

statePredByWindow = data.frame()
stateThresholdsByWindow = data.frame()

progress = txtProgressBar(max=length(rsvWindowYears), style=3)
idx = 0

for (y in rsvWindowYears) {

  stateYearObs = stateObsByWindow %>% filter(year==y)
  stateYearModel = gam(rsv ~ s(time, k=5, bs="cp", m=3), family=poisson, data=stateYearObs)

  stateYearPred = stateYearModel %>% 
    predict(type="response", newdata=modelTime, se.fit=TRUE)

  stateYearPred = modelTime %>% 
    cbind(data.frame(rsv.fit=stateYearPred$fit, rsv.fit.se=stateYearPred$se.fit)) %>%
    cbind(stateYearModel %>% outbreak.predict.timeseries(modelTime, function(model, params, newdata) { outbreak.calc.cum(model, params, newdata) })) %>%
    rename(rsv.cum.fit.lower=lower, rsv.cum.fit.upper=upper, rsv.cum.fit=median) %>%
    mutate(year=y)

  statePredByWindow = statePredByWindow %>% rbind(stateYearPred)

  stateYearThresholds = stateYearModel %>%
    outbreak.predict.scalars(
      modelTime,
      function(model, params, newdata) { outbreak.calc.thresholds(model, params, newdata$time, onset=seasonThreshold, offset=1 - seasonThreshold) }
    ) %>%
    mutate(year=y)

  stateThresholdsByWindow = stateThresholdsByWindow %>% rbind(stateYearThresholds)

  idx = idx + 1
  setTxtProgressBar(progress, idx)
}
```

```{r}
makeStateWindowPlots = function(c, plotter) {
  lapply(rsvWindowYears, function(y) {
    stateYearObs = stateObsByWindow %>% filter(year==y)
    stateYearPred = statePredByWindow %>% filter(year==y)
    stateYearThresholds = stateThresholdsByWindow %>% filter(year==y)
    plotter(y, stateYearObs, stateYearPred, stateYearThresholds)
  })
}

# Plot RSV cases with model prediction 
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    geom_ribbon(data=stateYearPred, aes(ymin = rsv.fit - rsv.fit.se, ymax = rsv.fit + rsv.fit.se), alpha = 0.2) +
    geom_line(data=stateYearPred, aes(y=rsv.fit)) +
    geom_hline(aes(yintercept = 0)) +
    geom_point(aes(y=rsv), shape=20) +
    annotate("rect",
             xmin=stateYearThresholds$onset.lower,
             xmax=stateYearThresholds$onset.upper,
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect",
             xmin=stateYearThresholds$offset.lower,
             xmax=stateYearThresholds$offset.upper,
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

# Plot RSV cumulative cases with model prediction 
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    geom_hline(aes(yintercept = 0)) +
    geom_point(shape=20, aes(y=rsv.cum)) +
    annotate("rect", 
             xmin=stateYearThresholds$onset.lower, 
             xmax=stateYearThresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=stateYearThresholds$offset.lower, 
             xmax=stateYearThresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

# Plot RSV cumulative case % with model prediction 
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    geom_hline(aes(yintercept = 0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=stateYearPred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=stateYearPred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=stateYearThresholds$onset.lower, 
             xmax=stateYearThresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=stateYearThresholds$offset.lower, 
             xmax=stateYearThresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

# Zoom in on onsets
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    coord_cartesian(
      xlim=c(min(stateYearThresholds$onset.lower) - 1, max(stateYearThresholds$onset.upper) + 1),
      ylim=c(0, 0.2)
    ) +
    geom_hline(aes(yintercept=0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=stateYearPred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=stateYearPred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=stateYearThresholds$onset.lower, 
             xmax=stateYearThresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=stateYearThresholds$offset.lower, 
             xmax=stateYearThresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

# Zoom in on offsets
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    coord_cartesian(
      xlim=c(min(stateYearThresholds$offset.lower) - 1, max(stateYearThresholds$offset.upper) + 1),
      ylim=c(0.8, 1)
    ) +
    geom_hline(aes(yintercept=0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=stateYearPred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=stateYearPred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=stateYearThresholds$onset.lower, 
             xmax=stateYearThresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=stateYearThresholds$offset.lower, 
             xmax=stateYearThresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

```

```{r}
onset = ggplot(stateThresholdsByWindow) +
  geom_errorbar(aes(ymin=onset.lower, ymax=onset.upper, x=year), width=.25) + 
  labs(x="RSV Season Onset")
offset = ggplot(stateThresholdsByWindow) +
  geom_errorbar(aes(ymin=offset.lower, ymax=offset.upper, x=year), width=.25) + 
  labs(x="RSV Season Offset")

ggplot2.multiplot(plotlist=list(onset, offset), cols=2)

```

```{r}
obsByCounty = dataset %>% 
  rename(time=weekofyear) %>% 
  group_by(county, time) %>% 
  summarize(rsv=sum(rsv)) %>% 
  left_join(rsvCountyGroups, by="county") %>%
  group_by(time) %>%
  do(summarizeCountyGroups(.)) %>%
  group_by(county) %>% 
  mutate(rsv.cum=cumsum(rsv), rsv.cum.frac=rsv.cum / sum(rsv)) %>% 
  as.data.frame()

predByCounty = data.frame()
thresholdsByCounty = data.frame()
fixedStratUnprotectedByCounty = data.frame()

progress = txtProgressBar(max=length(levels(obsByCounty$county)), style=3)
idx = 0

for (c in levels(obsByCounty$county)) {
  countyObs = obsByCounty %>% filter(county==c)
  countyModel = gam(rsv ~ s(time, k=20, bs="cp", m=3), family=poisson, data=countyObs)

  countyPred = countyModel %>% 
    predict(type="response", newdata=modelTime, se.fit=TRUE)

  countyPred = modelTime %>% 
    cbind(data.frame(rsv.fit=countyPred$fit, rsv.fit.se=countyPred$se.fit)) %>%
    cbind(countyModel %>% outbreak.predict.timeseries(modelTime, function(model, params, newdata) { outbreak.calc.cum(model, params, newdata) })) %>%
    rename(rsv.cum.fit.lower=lower, rsv.cum.fit.upper=upper, rsv.cum.fit=median) %>%
    mutate(county=factor(c, levels=levels(obsByCounty$county)))

  predByCounty = predByCounty %>% rbind(countyPred)

  countyThresholds = countyModel %>%
    outbreak.predict.scalars(
      modelTime,
      function(model, params, newdata) { 
        outbreak.calc.thresholds(model, params, newdata$time, onset=seasonThreshold, offset=1 - seasonThreshold)
      }
    ) %>%
    mutate(county=factor(c, levels=levels(obsByCounty$county)))

  thresholdsByCounty = thresholdsByCounty %>% rbind(countyThresholds)

  countyUnprotected = countyModel %>%
    outbreak.predict.scalars(
      modelTime,
      function(model, params, newdata) { 
        outbreak.calc.unprotected(model, params, newdata$time, ppxFixedStrategies(c))
      }
    ) %>%
    mutate(county=factor(c, levels=levels(obsByCounty$county)))

  fixedStratUnprotectedByCounty = fixedStratUnprotectedByCounty %>% rbind(countyUnprotected)

  idx = idx + 1
  setTxtProgressBar(progress, idx)
}

```

```{r}
makeCountyPlots = function(counties, plotter) {
  lapply(counties, function(c) {
    countyObs = obsByCounty %>% filter(county==c)
    countyPred = predByCounty %>% filter(county==c)
    countyThresholds = thresholdsByCounty %>% filter(county==c)
    plotter(c, countyObs, countyPred, countyThresholds)
  })
}

makeCountyYearPlots = function(c, plotter) {
  lapply(rsvYears, function(y) {
    countyYearObs = obsByCountyYear %>% filter(county==c, year==y)
    countyYearPred = predByCountyYear %>% filter(county==c, year==y)
    countyYearThresholds = thresholdsByCountyYear %>% filter(county==c, year==y)
    plotter(c, y, countyYearObs, countyYearPred, countyYearThresholds)
  })
}

# Plot RSV cases with model prediction 
ggplot_cases = function(obs, pred, thresholds, title) {
  ggplot(obs, aes(x=time)) +
    geom_ribbon(data=pred, aes(ymin = rsv.fit - rsv.fit.se, ymax = rsv.fit + rsv.fit.se), alpha = 0.2) +
    geom_line(data=pred, aes(y=rsv.fit)) +
    geom_hline(aes(yintercept = 0)) +
    geom_point(aes(y=rsv), shape=20) +
    annotate("rect",
             xmin=thresholds$onset.lower,
             xmax=thresholds$onset.upper,
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect",
             xmin=thresholds$offset.lower,
             xmax=thresholds$offset.upper,
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV cases, #", x="Week of year", title=title)
}

# Plot RSV cumulative cases (no model prediction)
ggplot_cum_cases = function(obs, pred, thresholds, title) {
  ggplot(obs, aes(x=time)) +
    geom_hline(aes(yintercept = 0)) +
    geom_point(shape=20, aes(y=rsv.cum)) +
    annotate("rect", 
             xmin=thresholds$onset.lower, 
             xmax=thresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=thresholds$offset.lower, 
             xmax=thresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV cumulative cases, #", x="Week of year", title=title)

}

# Plot RSV cumulative case percentage with model prediction
ggplot_cum_cases_pct = function(obs, pred, thresholds, title) {
  ggplot(obs, aes(x=time)) +
    geom_hline(aes(yintercept = 0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=pred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=pred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=thresholds$onset.lower, 
             xmax=thresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=thresholds$offset.lower, 
             xmax=thresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV cumulative cases, %", x="Week of year", title=title)
}

# Plot RSV cumulative case % with model prediction, zoomed in on onset
ggplot_cum_cases_pct_zoom = function(obs, pred, thresholds, xmin, xmax, ymin, ymax, title) {
  ggplot(obs, aes(x=time)) +
    coord_cartesian(
      xlim=c(xmin, xmax),
      ylim=c(ymin, ymax)
    ) +
    geom_hline(aes(yintercept = 0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=pred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=pred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=thresholds$onset.lower, 
             xmax=thresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=thresholds$offset.lower, 
             xmax=thresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV cumulative cases, %", x="Week of year", title=title)
}

ggplot_thresholds_over_time = function(thresholds, title) {
  onset = ggplot(thresholds) +
    geom_errorbar(aes(ymin=onset.lower, ymax=onset.upper, x=year), width=.25) + 
    labs(x="Year", y="RSV season onset week", title=title)
  offset = ggplot(thresholds) +
    geom_errorbar(aes(ymin=offset.lower, ymax=offset.upper, x=year), width=.25) + 
    labs(x="Year", y="RSV season offset week", title=title)

  ggplot2.multiplot(plotlist=list(onset, offset), cols=2)
}

ggplot_prophylaxis_strategy = function(obs, pred, protected, title) {
  ggplot(obs, aes(x=time)) +
    geom_ribbon(data=pred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=pred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=min(protected, na.rm=TRUE), xmax=max(protected, na.rm=TRUE), 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    geom_hline(aes(yintercept=0)) +
    labs(y="RSV", x="Time", title=title)
}

```

```{r}
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cases(countyObs, countyPred, countyThresholds, c)
}), cols=4)

# Plot RSV cumulative cases with model prediction 
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cum_cases(countyObs, countyPred, countyThresholds, c)
}), cols=4)

# Plot RSV cumulative case % with model prediction 
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cum_cases_pct(countyObs, countyPred, countyThresholds, c)
}), cols=4)

# Zoom in on onsets
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cum_cases_pct_zoom(countyObs, countyPred, countyThresholds, xmin=min(countyThresholds$onset.lower) - 1, xmax=max(countyThresholds$onset.upper) + 1, ymin=0, ymax=0.2, c)
}), cols=4)

# Zoom in on offsets
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cum_cases_pct_zoom(countyObs, countyPred, countyThresholds, xmin=min(countyThresholds$offset.lower) - 1, xmax=max(countyThresholds$offset.upper) + 1, ymin=0.8, ymax=1, c)
}), cols=4)
```

```{r}
onset = ggplot(thresholdsByCounty) +
  geom_errorbarh(aes(x=onset.median, xmin=onset.lower, xmax=onset.upper, y=county), height=.25) + 
  labs(x="RSV Season Onset")
offset = ggplot(thresholdsByCounty) +
  geom_errorbarh(aes(x=offset.median, xmin=offset.lower, xmax=offset.upper, y=county), height=.25) + 
  labs(x="RSV Season Offset")

ggplot2.multiplot(plotlist=list(onset, offset), cols=2)

```

```{r warning=FALSE}
obsByCountyYear = dataset %>% 
  rename(time=weekofyear) %>% 
  group_by(county, weeki, year, time) %>% 
  summarize(rsv=sum(rsv)) %>% 
  left_join(rsvCountyGroups, by="county") %>%
  group_by(weeki, year, time) %>%
  do(summarizeCountyGroups(.)) %>%
  group_by(county, year) %>%
  mutate(rsv.cum=cumsum(rsv), rsv.cum.frac=rsv.cum / if_else(sum(rsv) > 0, sum(rsv), as.integer(1))) %>% 
  select(-weeki) %>%
  as.data.frame()

predByCountyYear = data.frame()
thresholdsByCountyYear = data.frame()
slidingStratUnprotectedSimByCountyYear = data.frame()

progress = txtProgressBar(max=length(levels(obsByCounty$county)) * length(rsvYears) * 2, style=3)
idx = 0

for (c in levels(obsByCountyYear$county)) {
  for (y in rsvYears) {
    countyYearObs = obsByCountyYear %>% filter(county==c, year==y)
    countyYearModel = gam(rsv ~ s(time, k=5, bs="cp", m=3), family=poisson, data=countyYearObs)

    countyYearPred = countyYearModel %>% 
      predict(type="response", newdata=modelTime, se.fit=TRUE)

    countyYearPred = modelTime %>% 
      cbind(data.frame(rsv.fit=countyYearPred$fit, rsv.fit.se=countyYearPred$se.fit)) %>%
      cbind(countyYearModel %>% outbreak.predict.timeseries(modelTime, function(model, params, newdata) { outbreak.calc.cum(model, params, newdata) })) %>%
      rename(rsv.cum.fit.lower=lower, rsv.cum.fit.upper=upper, rsv.cum.fit=median) %>%
      mutate(county=factor(c, levels=levels(obsByCountyYear$county)), year=y)

    predByCountyYear = predByCountyYear %>% rbind(countyYearPred)

    countyYearThresholds = countyYearModel %>%
      outbreak.predict.scalars(
        modelTime,
        function(model, params, newdata) { outbreak.calc.thresholds(model, params, newdata$time, onset=seasonThreshold, offset=1 - seasonThreshold) }
      ) %>%
      mutate(county=factor(c, levels=levels(obsByCountyYear$county)), year=y)

    thresholdsByCountyYear = thresholdsByCountyYear %>% rbind(countyYearThresholds)

    if (y %in% rsvWindowYears) {
      countyYearUnprotectedSim = countyYearModel %>%
        outbreak.predict.scalars.sim(
          modelTime,
          function(model, params, newdata) { 
            outbreak.calc.unprotected(model, params, newdata$time, ppxSlidingStrategies(c, y))
          }
        ) %>%
        mutate(county=factor(c, levels=levels(obsByCountyYear$county)), year=y)

      slidingStratUnprotectedSimByCountyYear = slidingStratUnprotectedSimByCountyYear %>% rbind(countyYearUnprotectedSim)
    }

    idx = idx + 1
    setTxtProgressBar(progress, idx)
  }
}

# Aggregate sliding strategies over the entire time span

unprotectedByCounty = slidingStratUnprotectedSimByCountyYear %>% 
  group_by(county, outbreak.sim) %>%
  summarize_at(vars(contains("total"), contains("count")), sum) %>%
  group_by(county, outbreak.sim) %>%
  do((function(df) {
    stratNames = names(ppxSlidingStrategies())
    for (stratName in stratNames) {
      countName = sprintf("%s.count", stratName)
      fracName = sprintf("%s.frac", stratName)
      df = df %>% 
        mutate_at(fracName, function(.) { df[[countName]] / df$total })
    }
    df
  }) (.)) %>%
  group_by(county) %>%
  do((function(df) {
    df %>%
      select(-county, -total, -matches(".count")) %>%
      outbreak.predict.scalars.confints () %>%
      mutate(county=unique(df$county))
  })(.)) %>%
  inner_join(fixedStratUnprotectedByCounty, by="county")
```

```{r}
for (c in levels(obsByCountyYear$county)) {
  countyThresholdsByYear = thresholdsByCountyYear %>% filter(county==c)

  ggplot_thresholds_over_time(countyThresholdsByYear, title=c)
}

```

```{r}
# Plot fixed propxylaxis strategies
for (c in levels(obsByCounty$county)) {
  d = "plots/by county"
  if (!dir.exists(d)) {
    dir.create(d, showWarnings=FALSE, recursive=TRUE)
  }

  countyObs = obsByCounty %>% filter(county==c)
  countyPred = predByCounty %>% filter(county==c)

  ggsave(sprintf("%s/%s prophylaxis strategies.pdf", d, c), ggplot2.multiplot(plotlist=lapply(names(ppxFixedStrategies(c)), function(stratName) {
    strat = ppxFixedStrategies(c)[[stratName]]
    protected = modelTime$time %>% sapply(function(t) { if (strat(t) == 0) NA else t })

    ggplot_prophylaxis_strategy(countyObs, countyPred, protected, sprintf("%s w/ %s", c, stratName)) + labs(x=NULL)
  }), cols=1), width=4, height=6, units="in")
}

# Plot sliding propxylaxis strategies
for (c in levels(obsByCounty$county)) {
  for (y in rsvWindowYears) {
    d = sprintf("plots/by county and year/%s", c)
    if (!dir.exists(d)) {
      dir.create(d, showWarnings=FALSE, recursive=TRUE)
    }

    countyYearObs = obsByCountyYear %>% filter(county==c, year==y)
    countyYearPred = predByCountyYear %>% filter(county==c, year==y)

    ggsave(sprintf("%s/%s prophylaxis strategies.pdf", d, y), ggplot2.multiplot(plotlist=lapply(names(ppxSlidingStrategies(c, y)), function(stratName) {
      strat = ppxSlidingStrategies(c, y)[[stratName]]
      protected = modelTime$time %>% sapply(function(t) { if (strat(t) == 0) NA else t })

      ggplot_prophylaxis_strategy(countyYearObs, countyYearPred, protected, sprintf("%s w/ %s", c, stratName)) + labs(x=NULL)
    }), cols=1), width=4, height=6, units="in")
  }
}
```

```{r}
stratNames = c(names(ppxFixedStrategies()), names(ppxSlidingStrategies()))
plot = ggplot(unprotectedByCounty) +
  scale_linetype_manual(breaks=stratNames, values=seq_along(stratNames)) + 
  coord_cartesian(
    xlim=c(0, 0.1)
  ) +
  labs(x="Unprotected cases", y=NULL) +
  theme(aspect.ratio=2, panel.grid.major.y=element_blank())

for (idx in seq_along(stratNames)) {
  stratName = stratNames[idx]
  median = sprintf("%s.frac.median", stratName)
  lower = sprintf("%s.frac.lower", stratName)
  upper = sprintf("%s.frac.upper", stratName)
  linetype = sprintf('"%s"', stratName)

  delta = 0.1
  nudge = (idx - (length(stratNames) + 1) / 2) * -delta

  plot = plot + 
    geom_errorbarh(aes_string(x=median, xmin=lower, xmax=upper, y="county", linetype=linetype), height=0, position=position_nudge(y=nudge)) + 
    geom_point(aes_string(x=lower, y="county"), shape=20, position=position_nudge(y=nudge)) +
    geom_point(aes_string(x=upper, y="county"), shape=20, position=position_nudge(y=nudge))
}

plot
ggsave("plots/prophylaxis strategy comparison.pdf", plot, height=8, width=12, units="in")
```

```{r}
ignore_ = lapply(levels(obsByCountyYear$county), function(c) {
  d = sprintf("plots/by county and year/%s", c)
  if (!dir.exists(d)) {
    dir.create(d, showWarnings=FALSE, recursive=TRUE)
  }

  multiplot = ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cases(countyYearObs, countyYearPred, countyYearThresholds, NULL)
    ggsave(sprintf("%s/%s cases.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)

  # Plot RSV cumulative cases with model prediction 
  ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cum_cases(countyYearObs, countyYearPred, countyYearThresholds, NULL)
    ggsave(sprintf("%s/%s cum cases.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)

  # Plot RSV cumulative case % with model prediction 
  ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cum_cases_pct(countyYearObs, countyYearPred, countyYearThresholds, NULL)
    ggsave(sprintf("%s/%s model.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)

  # Zoom in on onsets
  ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cum_cases_pct_zoom(countyYearObs, countyYearPred, countyYearThresholds, xmin=min(countyYearThresholds$onset.lower) - 1, xmax=max(countyYearThresholds$onset.upper) + 1, ymin=0, ymax=0.2, NULL)
    ggsave(sprintf("%s/%s onset.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)

  # Zoom in on offsets
  ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cum_cases_pct_zoom(countyYearObs, countyYearPred, countyYearThresholds, xmin=min(countyYearThresholds$offset.lower) - 1, xmax=max(countyYearThresholds$offset.upper) + 1, ymin=0.8, ymax=1, c)
    ggsave(sprintf("%s/%s offset.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)
})
```