### 2018-01-02 Analysis/RSV.Rmd
Refactoring code for paper into separate R file

---
title: "RSV"
output: html_notebook
---

```{r setup, include=FALSE, error=FALSE, warning=FALSE, message=FALSE}
install.packages(c("dplyr", "zipcode", "reshape", "mgcv", "ggplot2", "gridExtra", "devtools"))
library(devtools)
install_github("gavinsimpson/tsgam")
install_github("kassambara/easyGgplot2")
install_github("airbornemint/outbreak-predict")

source("./RSV.R")
```

```{r}
ggplot(dataset %>% 
         group_by(weeki) %>%
         summarize(rsv=sum(rsv)), 
       aes(x=weeki, y=rsv)) +
  geom_hline(aes(yintercept = 0)) +
  geom_line() +
  labs(y="RSV cases/week", x="Time", title="All counties")

plots = lapply(rsvCounties, function(c) {
  ggplot(dataset %>% 
           filter(county==c) %>%
           group_by(weeki) %>%
           summarize(rsv=sum(rsv)), 
         aes(x=weeki, y=rsv)) +
    geom_hline(aes(yintercept = 0)) +
    geom_line() +
    labs(y="RSV cases/week", x="Time", title=c)
})

ggplot2.multiplot(plotlist=plots, cols=4)

```

```{r}
ppxSlidingStrategies = function(c=NULL, y=NULL) {
  if (!is.null(y) && !is.null(c)) {
    stateSlidingOnset = (stateThresholdsByWindow %>% filter(year==y))$onset.median
    stateSlidingOffset = (stateThresholdsByWindow %>% filter(year==y))$offset.median
  }

  list(
    stateSlidingOnset=function(time) {
      evalOnsetStrategy(stateSlidingOnset, time)
    },
    stateSlidingOffset=function(time) {
      evalOffsetStrategy(stateSlidingOffset, time)
    }
  )
}

slidingWindowDuration = 3 # Years

```

```{r}
stateObsByWindow = dataset %>% 
  mutate(time=weeki) %>% 
  group_by(time, year, weekofyear) %>% 
  summarize(rsv=sum(rsv)) %>%
  as.data.frame()

total = data.frame(time=stateObsByWindow$time, rsv=0)

for (idx in seq(1, slidingWindowDuration)) {
  total = total %>%
    inner_join(stateObsByWindow %>%
                 mutate(time=time + 52 * idx) %>%
                 select(time, rsv),
               by="time"
               ) %>%
    mutate(rsv=rsv.x + rsv.y) %>%
    select(time, rsv)
}

stateObsByWindow = stateObsByWindow %>%
  select(-rsv) %>%
  inner_join(total, by="time") %>%
  group_by(year) %>%
  mutate(rsv.cum=cumsum(rsv), rsv.cum.frac=rsv.cum / sum(rsv)) %>% 
  select(-time) %>%
  rename(time=weekofyear) %>%
  filter(year >= min(rsvYears) + slidingWindowDuration) %>%
  as.data.frame()

rsvWindowYears = unique(stateObsByWindow$year)

statePredByWindow = data.frame()
stateThresholdsByWindow = data.frame()

progress = txtProgressBar(max=length(rsvWindowYears), style=3)
idx = 0

for (y in rsvWindowYears) {

  stateYearObs = stateObsByWindow %>% filter(year==y)
  stateYearModel = gam(rsv ~ s(time, k=5, bs="cp", m=3), family=poisson, data=stateYearObs)

  stateYearPred = stateYearModel %>% 
    predict(type="response", newdata=modelTime, se.fit=TRUE)

  stateYearPred = modelTime %>% 
    cbind(data.frame(rsv.fit=stateYearPred$fit, rsv.fit.se=stateYearPred$se.fit)) %>%
    cbind(stateYearModel %>% outbreak.predict.timeseries(modelTime, function(model, params, newdata) { outbreak.calc.cum(model, params, newdata) })) %>%
    rename(rsv.cum.fit.lower=lower, rsv.cum.fit.upper=upper, rsv.cum.fit=median) %>%
    mutate(year=y)

  statePredByWindow = statePredByWindow %>% rbind(stateYearPred)

  stateYearThresholds = stateYearModel %>%
    outbreak.predict.scalars(
      modelTime,
      function(model, params, newdata) { outbreak.calc.thresholds(model, params, newdata$time, onset=seasonThreshold, offset=1 - seasonThreshold) }
    ) %>%
    mutate(year=y)

  stateThresholdsByWindow = stateThresholdsByWindow %>% rbind(stateYearThresholds)

  idx = idx + 1
  setTxtProgressBar(progress, idx)
}
```

```{r}
makeStateWindowPlots = function(c, plotter) {
  lapply(rsvWindowYears, function(y) {
    stateYearObs = stateObsByWindow %>% filter(year==y)
    stateYearPred = statePredByWindow %>% filter(year==y)
    stateYearThresholds = stateThresholdsByWindow %>% filter(year==y)
    plotter(y, stateYearObs, stateYearPred, stateYearThresholds)
  })
}

# Plot RSV cases with model prediction 
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    geom_ribbon(data=stateYearPred, aes(ymin = rsv.fit - rsv.fit.se, ymax = rsv.fit + rsv.fit.se), alpha = 0.2) +
    geom_line(data=stateYearPred, aes(y=rsv.fit)) +
    geom_hline(aes(yintercept = 0)) +
    geom_point(aes(y=rsv), shape=20) +
    annotate("rect",
             xmin=stateYearThresholds$onset.lower,
             xmax=stateYearThresholds$onset.upper,
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect",
             xmin=stateYearThresholds$offset.lower,
             xmax=stateYearThresholds$offset.upper,
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

# Plot RSV cumulative cases with model prediction 
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    geom_hline(aes(yintercept = 0)) +
    geom_point(shape=20, aes(y=rsv.cum)) +
    annotate("rect", 
             xmin=stateYearThresholds$onset.lower, 
             xmax=stateYearThresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=stateYearThresholds$offset.lower, 
             xmax=stateYearThresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

# Plot RSV cumulative case % with model prediction 
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    geom_hline(aes(yintercept = 0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=stateYearPred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=stateYearPred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=stateYearThresholds$onset.lower, 
             xmax=stateYearThresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=stateYearThresholds$offset.lower, 
             xmax=stateYearThresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

# Zoom in on onsets
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    coord_cartesian(
      xlim=c(min(stateYearThresholds$onset.lower) - 1, max(stateYearThresholds$onset.upper) + 1),
      ylim=c(0, 0.2)
    ) +
    geom_hline(aes(yintercept=0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=stateYearPred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=stateYearPred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=stateYearThresholds$onset.lower, 
             xmax=stateYearThresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=stateYearThresholds$offset.lower, 
             xmax=stateYearThresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

# Zoom in on offsets
ggplot2.multiplot(plotlist=makeStateWindowPlots(rsvWindowYears, function(y, stateYearObs, stateYearPred, stateYearThresholds) {
  ggplot(stateYearObs, aes(x=time)) +
    coord_cartesian(
      xlim=c(min(stateYearThresholds$offset.lower) - 1, max(stateYearThresholds$offset.upper) + 1),
      ylim=c(0.8, 1)
    ) +
    geom_hline(aes(yintercept=0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=stateYearPred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=stateYearPred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=stateYearThresholds$onset.lower, 
             xmax=stateYearThresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=stateYearThresholds$offset.lower, 
             xmax=stateYearThresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV", x="Time", title=y)
}), cols=4)

```

```{r}
onset = ggplot(stateThresholdsByWindow) +
  geom_errorbar(aes(ymin=onset.lower, ymax=onset.upper, x=year), width=.25) + 
  labs(x="RSV Season Onset")
offset = ggplot(stateThresholdsByWindow) +
  geom_errorbar(aes(ymin=offset.lower, ymax=offset.upper, x=year), width=.25) + 
  labs(x="RSV Season Offset")

ggplot2.multiplot(plotlist=list(onset, offset), cols=2)

```

```{r}
makeCountyPlots = function(counties, plotter) {
  lapply(counties, function(c) {
    countyObs = obsByCounty %>% filter(county==c)
    countyPred = predByCounty %>% filter(county==c)
    countyThresholds = thresholdsByCounty %>% filter(county==c)
    plotter(c, countyObs, countyPred, countyThresholds)
  })
}

makeCountyYearPlots = function(c, plotter) {
  lapply(rsvYears, function(y) {
    countyYearObs = obsByCountyYear %>% filter(county==c, year==y)
    countyYearPred = predByCountyYear %>% filter(county==c, year==y)
    countyYearThresholds = thresholdsByCountyYear %>% filter(county==c, year==y)
    plotter(c, y, countyYearObs, countyYearPred, countyYearThresholds)
  })
}

# Plot RSV cases with model prediction 
ggplot_cases = function(obs, pred, thresholds, title) {
  ggplot(obs, aes(x=time)) +
    geom_ribbon(data=pred, aes(ymin = rsv.fit - rsv.fit.se, ymax = rsv.fit + rsv.fit.se), alpha = 0.2) +
    geom_line(data=pred, aes(y=rsv.fit)) +
    geom_hline(aes(yintercept = 0)) +
    geom_point(aes(y=rsv), shape=20) +
    annotate("rect",
             xmin=thresholds$onset.lower,
             xmax=thresholds$onset.upper,
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect",
             xmin=thresholds$offset.lower,
             xmax=thresholds$offset.upper,
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV cases, #", x="Week of year", title=title)
}

# Plot RSV cumulative cases (no model prediction)
ggplot_cum_cases = function(obs, pred, thresholds, title) {
  ggplot(obs, aes(x=time)) +
    geom_hline(aes(yintercept = 0)) +
    geom_point(shape=20, aes(y=rsv.cum)) +
    annotate("rect", 
             xmin=thresholds$onset.lower, 
             xmax=thresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=thresholds$offset.lower, 
             xmax=thresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV cumulative cases, #", x="Week of year", title=title)

}

# Plot RSV cumulative case percentage with model prediction
ggplot_cum_cases_pct = function(obs, pred, thresholds, title) {
  ggplot(obs, aes(x=time)) +
    geom_hline(aes(yintercept = 0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=pred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=pred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=thresholds$onset.lower, 
             xmax=thresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=thresholds$offset.lower, 
             xmax=thresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV cumulative cases, %", x="Week of year", title=title)
}

# Plot RSV cumulative case % with model prediction, zoomed in on onset
ggplot_cum_cases_pct_zoom = function(obs, pred, thresholds, xmin, xmax, ymin, ymax, title) {
  ggplot(obs, aes(x=time)) +
    coord_cartesian(
      xlim=c(xmin, xmax),
      ylim=c(ymin, ymax)
    ) +
    geom_hline(aes(yintercept = 0)) +
    geom_hline(aes(yintercept=seasonThreshold), color=gray(.5)) +
    geom_hline(aes(yintercept=1 - seasonThreshold), color=gray(.5)) +
    geom_ribbon(data=pred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=pred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=thresholds$onset.lower, 
             xmax=thresholds$onset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    annotate("rect", 
             xmin=thresholds$offset.lower, 
             xmax=thresholds$offset.upper, 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    labs(y="RSV cumulative cases, %", x="Week of year", title=title)
}

ggplot_thresholds_over_time = function(thresholds, title) {
  onset = ggplot(thresholds) +
    geom_errorbar(aes(ymin=onset.lower, ymax=onset.upper, x=year), width=.25) + 
    labs(x="Year", y="RSV season onset week", title=title)
  offset = ggplot(thresholds) +
    geom_errorbar(aes(ymin=offset.lower, ymax=offset.upper, x=year), width=.25) + 
    labs(x="Year", y="RSV season offset week", title=title)

  ggplot2.multiplot(plotlist=list(onset, offset), cols=2)
}

ggplot_prophylaxis_strategy = function(obs, pred, protected, title) {
  ggplot(obs, aes(x=time)) +
    geom_ribbon(data=pred, aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper), alpha=0.2) +
    geom_line(data=pred, aes(y=rsv.cum.fit)) +
    geom_point(shape=20, aes(y=rsv.cum.frac)) +
    annotate("rect", 
             xmin=min(protected, na.rm=TRUE), xmax=max(protected, na.rm=TRUE), 
             ymin=-Inf, ymax=Inf, alpha=.25) +
    geom_hline(aes(yintercept=0)) +
    labs(y="RSV", x="Time", title=title)
}

```

```{r}
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cases(countyObs, countyPred, countyThresholds, c)
}), cols=4)

# Plot RSV cumulative cases with model prediction 
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cum_cases(countyObs, countyPred, countyThresholds, c)
}), cols=4)

# Plot RSV cumulative case % with model prediction 
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cum_cases_pct(countyObs, countyPred, countyThresholds, c)
}), cols=4)

# Zoom in on onsets
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cum_cases_pct_zoom(countyObs, countyPred, countyThresholds, xmin=min(countyThresholds$onset.lower) - 1, xmax=max(countyThresholds$onset.upper) + 1, ymin=0, ymax=0.2, c)
}), cols=4)

# Zoom in on offsets
ggplot2.multiplot(plotlist=makeCountyPlots(levels(obsByCounty$county), function(c, countyObs, countyPred, countyThresholds) {
  ggplot_cum_cases_pct_zoom(countyObs, countyPred, countyThresholds, xmin=min(countyThresholds$offset.lower) - 1, xmax=max(countyThresholds$offset.upper) + 1, ymin=0.8, ymax=1, c)
}), cols=4)
```

```{r}
onset = ggplot(thresholdsByCounty) +
  geom_errorbarh(aes(x=onset.median, xmin=onset.lower, xmax=onset.upper, y=county), height=.25) + 
  labs(x="RSV Season Onset")
offset = ggplot(thresholdsByCounty) +
  geom_errorbarh(aes(x=offset.median, xmin=offset.lower, xmax=offset.upper, y=county), height=.25) + 
  labs(x="RSV Season Offset")

ggplot2.multiplot(plotlist=list(onset, offset), cols=2)

```

```{r warning=FALSE}
obsByCountyYear = dataset %>% 
  rename(time=weekofyear) %>% 
  group_by(county, weeki, year, time) %>% 
  summarize(rsv=sum(rsv)) %>% 
  left_join(rsvCountyGroups, by="county") %>%
  group_by(weeki, year, time) %>%
  do(summarizeCountyGroups(.)) %>%
  group_by(county, year) %>%
  mutate(rsv.cum=cumsum(rsv), rsv.cum.frac=rsv.cum / if_else(sum(rsv) > 0, sum(rsv), as.integer(1))) %>% 
  select(-weeki) %>%
  as.data.frame()

predByCountyYear = data.frame()
thresholdsByCountyYear = data.frame()
slidingStratUnprotectedSimByCountyYear = data.frame()

progress = txtProgressBar(max=length(levels(obsByCounty$county)) * length(rsvYears) * 2, style=3)
idx = 0

for (c in levels(obsByCountyYear$county)) {
  for (y in rsvYears) {
    countyYearObs = obsByCountyYear %>% filter(county==c, year==y)
    countyYearModel = gam(rsv ~ s(time, k=5, bs="cp", m=3), family=poisson, data=countyYearObs)

    countyYearPred = countyYearModel %>% 
      predict(type="response", newdata=modelTime, se.fit=TRUE)

    countyYearPred = modelTime %>% 
      cbind(data.frame(rsv.fit=countyYearPred$fit, rsv.fit.se=countyYearPred$se.fit)) %>%
      cbind(countyYearModel %>% outbreak.predict.timeseries(modelTime, function(model, params, newdata) { outbreak.calc.cum(model, params, newdata) })) %>%
      rename(rsv.cum.fit.lower=lower, rsv.cum.fit.upper=upper, rsv.cum.fit=median) %>%
      mutate(county=factor(c, levels=levels(obsByCountyYear$county)), year=y)

    predByCountyYear = predByCountyYear %>% rbind(countyYearPred)

    countyYearThresholds = countyYearModel %>%
      outbreak.predict.scalars(
        modelTime,
        function(model, params, newdata) { outbreak.calc.thresholds(model, params, newdata$time, onset=seasonThreshold, offset=1 - seasonThreshold) }
      ) %>%
      mutate(county=factor(c, levels=levels(obsByCountyYear$county)), year=y)

    thresholdsByCountyYear = thresholdsByCountyYear %>% rbind(countyYearThresholds)

    if (y %in% rsvWindowYears) {
      countyYearUnprotectedSim = countyYearModel %>%
        outbreak.predict.scalars.sim(
          modelTime,
          function(model, params, newdata) { 
            outbreak.calc.unprotected(model, params, newdata$time, ppxSlidingStrategies(c, y))
          }
        ) %>%
        mutate(county=factor(c, levels=levels(obsByCountyYear$county)), year=y)

      slidingStratUnprotectedSimByCountyYear = slidingStratUnprotectedSimByCountyYear %>% rbind(countyYearUnprotectedSim)
    }

    idx = idx + 1
    setTxtProgressBar(progress, idx)
  }
}

# Aggregate sliding strategies over the entire time span

unprotectedByCounty = slidingStratUnprotectedSimByCountyYear %>% 
  group_by(county, outbreak.sim) %>%
  summarize_at(vars(contains("total"), contains("count")), sum) %>%
  group_by(county, outbreak.sim) %>%
  do((function(df) {
    stratNames = names(ppxSlidingStrategies())
    for (stratName in stratNames) {
      countName = sprintf("%s.count", stratName)
      fracName = sprintf("%s.frac", stratName)
      df = df %>% 
        mutate_at(fracName, function(.) { df[[countName]] / df$total })
    }
    df
  }) (.)) %>%
  group_by(county) %>%
  do((function(df) {
    df %>%
      select(-county, -total, -matches(".count")) %>%
      outbreak.predict.scalars.confints () %>%
      mutate(county=unique(df$county))
  })(.)) %>%
  inner_join(fixedStratUnprotectedByCounty, by="county")
```

```{r}
for (c in levels(obsByCountyYear$county)) {
  countyThresholdsByYear = thresholdsByCountyYear %>% filter(county==c)

  ggplot_thresholds_over_time(countyThresholdsByYear, title=c)
}

```

```{r}
# Plot fixed propxylaxis strategies
for (c in levels(obsByCounty$county)) {
  d = "plots/by county"
  if (!dir.exists(d)) {
    dir.create(d, showWarnings=FALSE, recursive=TRUE)
  }

  countyObs = obsByCounty %>% filter(county==c)
  countyPred = predByCounty %>% filter(county==c)

  ggsave(sprintf("%s/%s prophylaxis strategies.pdf", d, c), ggplot2.multiplot(plotlist=lapply(names(ppxFixedStrategies(c)), function(stratName) {
    strat = ppxFixedStrategies(c)[[stratName]]
    protected = modelTime$time %>% sapply(function(t) { if (strat(t) == 0) NA else t })

    ggplot_prophylaxis_strategy(countyObs, countyPred, protected, sprintf("%s w/ %s", c, stratName)) + labs(x=NULL)
  }), cols=1), width=4, height=6, units="in")
}

# Plot sliding propxylaxis strategies
for (c in levels(obsByCounty$county)) {
  for (y in rsvWindowYears) {
    d = sprintf("plots/by county and year/%s", c)
    if (!dir.exists(d)) {
      dir.create(d, showWarnings=FALSE, recursive=TRUE)
    }

    countyYearObs = obsByCountyYear %>% filter(county==c, year==y)
    countyYearPred = predByCountyYear %>% filter(county==c, year==y)

    ggsave(sprintf("%s/%s prophylaxis strategies.pdf", d, y), ggplot2.multiplot(plotlist=lapply(names(ppxSlidingStrategies(c, y)), function(stratName) {
      strat = ppxSlidingStrategies(c, y)[[stratName]]
      protected = modelTime$time %>% sapply(function(t) { if (strat(t) == 0) NA else t })

      ggplot_prophylaxis_strategy(countyYearObs, countyYearPred, protected, sprintf("%s w/ %s", c, stratName)) + labs(x=NULL)
    }), cols=1), width=4, height=6, units="in")
  }
}
```

```{r}
stratNames = c(names(ppxFixedStrategies()), names(ppxSlidingStrategies()))
plot = ggplot(unprotectedByCounty) +
  scale_linetype_manual(breaks=stratNames, values=seq_along(stratNames)) + 
  coord_cartesian(
    xlim=c(0, 0.1)
  ) +
  labs(x="Unprotected cases", y=NULL) +
  theme(aspect.ratio=2, panel.grid.major.y=element_blank())

for (idx in seq_along(stratNames)) {
  stratName = stratNames[idx]
  median = sprintf("%s.frac.median", stratName)
  lower = sprintf("%s.frac.lower", stratName)
  upper = sprintf("%s.frac.upper", stratName)
  linetype = sprintf('"%s"', stratName)

  delta = 0.1
  nudge = (idx - (length(stratNames) + 1) / 2) * -delta

  plot = plot + 
    geom_errorbarh(aes_string(x=median, xmin=lower, xmax=upper, y="county", linetype=linetype), height=0, position=position_nudge(y=nudge)) + 
    geom_point(aes_string(x=lower, y="county"), shape=20, position=position_nudge(y=nudge)) +
    geom_point(aes_string(x=upper, y="county"), shape=20, position=position_nudge(y=nudge))
}

plot
ggsave("plots/prophylaxis strategy comparison.pdf", plot, height=8, width=12, units="in")
```

```{r}
ignore_ = lapply(levels(obsByCountyYear$county), function(c) {
  d = sprintf("plots/by county and year/%s", c)
  if (!dir.exists(d)) {
    dir.create(d, showWarnings=FALSE, recursive=TRUE)
  }

  multiplot = ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cases(countyYearObs, countyYearPred, countyYearThresholds, NULL)
    ggsave(sprintf("%s/%s cases.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)

  # Plot RSV cumulative cases with model prediction 
  ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cum_cases(countyYearObs, countyYearPred, countyYearThresholds, NULL)
    ggsave(sprintf("%s/%s cum cases.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)

  # Plot RSV cumulative case % with model prediction 
  ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cum_cases_pct(countyYearObs, countyYearPred, countyYearThresholds, NULL)
    ggsave(sprintf("%s/%s model.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)

  # Zoom in on onsets
  ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cum_cases_pct_zoom(countyYearObs, countyYearPred, countyYearThresholds, xmin=min(countyYearThresholds$onset.lower) - 1, xmax=max(countyYearThresholds$onset.upper) + 1, ymin=0, ymax=0.2, NULL)
    ggsave(sprintf("%s/%s onset.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)

  # Zoom in on offsets
  ggplot2.multiplot(plotlist=makeCountyYearPlots(c, function(c, y, countyYearObs, countyYearPred, countyYearThresholds) {
    plot = ggplot_cum_cases_pct_zoom(countyYearObs, countyYearPred, countyYearThresholds, xmin=min(countyYearThresholds$offset.lower) - 1, xmax=max(countyYearThresholds$offset.upper) + 1, ymin=0.8, ymax=1, c)
    ggsave(sprintf("%s/%s offset.pdf", d, y), height=4, width=6, units="in")
    plot
  }), cols=4)
})
```