### 2018-04-30 Thesis/Poster.rnw
Added YSPH research day poster

<<options, echo=FALSE>>=
    opts_chunk$set(echo=FALSE, message=FALSE)
@

<<load>>=
read_chunk("../Analysis/RSV.R")
read_chunk("../Thesis/Thesis.R")
@

<<libraries, cache=FALSE>>=
@

<<external, cache=TRUE>>=
@

<<util, warning=FALSE>>=
@

<<posterUtil, cache=TRUE>>=
posterPlotWidth = 19
posterPlotHeight = posterPlotWidth * 3 / 10
plotTextBaseSize = 24
@

<<coverageByRegionAAPInsightPoster, results='hide'>>=
    tikz(sprintf("%s/coverageByRegionAAPInsightPoster.tex", figuresDir), width=posterPlotWidth, height=posterPlotHeight, pointsize=37)

    strat = ppxFixedStrategies("all")[["aap"]]

    data = predByCounty %>%
	    filter(county %in% countiesForPlots)

    thresholds = data %>%
    	mutate(ppx=strat(time)) %>%
    	filter(ppx > 0) %>%
    	group_by(county) %>%
    	do(
    		data.frame(
    			ppx.start=min(.$time),
    			ppx.end=max(.$time),
    			unprotected.start=min(.$rsv.cum.fit),
    			unprotected.end=max(.$rsv.cum.fit)
    		)
    	) %>%
    	ungroup() %>%
    	as.data.frame() %>%
    	inner_join(thresholdsByCounty, by="county")

    barYmin = 0.025
    barYmax = 0.075

    spanExtents = data.frame(
        variable=c("ppx", "season"), 
        barYmin=c(1 + barYmin, -barYmax), 
        barYmax=c(1 + barYmax, -barYmin), 
        lineYmin=c(0, -barYmin), 
        lineYmax=c(1 + barYmin, 1)
    )

    spans = thresholds %>%
        select(county, matches("ppx."), matches(".median")) %>%
        rename(season.start=onset.median, season.end=offset.median) %>%
        melt(id.vars="county") %>%
        cSplit("variable", ".") %>%
        rename(variable=variable_1, part=variable_2) %>%
        cast(county + variable ~ part) %>%
        right_join(spanExtents, by="variable")

    spanLabels = legendLabels(c("AAP-recommended prophylaxis", "RSV season"), 20)

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_ribbon(aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper, x=time), fill=grey(.25)) +
    	geom_segment(aes(x=start, y=lineYmin, xend=start, yend=lineYmax, color=variable), data=spans, size=2) + 
    	geom_segment(aes(x=end, y=lineYmin, xend=end, yend=lineYmax, color=variable), data=spans, size=2) + 
    	geom_rect(aes(xmin=start, ymin=barYmin, xmax=end, ymax=barYmax, fill=variable, color=variable), data=spans, size=2) + 
    	geom_segment(aes(x=-Inf, y=unprotected.start, xend=ppx.start, yend=unprotected.start), linetype="55", data=thresholds, size=2) + 
        geom_segment(aes(x=-Inf, y=unprotected.end, xend=ppx.end, yend=unprotected.end), linetype="55", data=thresholds, size=2) + 
        geom_point(aes(x=min(monthBoundaries$min), y=unprotected.end), data=thresholds, shape=18) + 
        geom_point(aes(x=min(monthBoundaries$min), y=unprotected.start), data=thresholds, shape=18) + 
        scale_y_continuous(labels=latexPercent, breaks=seq(0, 1, by=.2), limits=c(-0.1, 1.1), expand=c(0, 0)) +
        scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
        scale_color_manual(labels=spanLabels, values=c(darkPurple, darkRed)) +
        scale_fill_manual(labels=spanLabels, values=c(darkPurple, darkRed)) +
        guides(color=guide_legend(keywidth=3, keyheight=.4, label.vjust=0.2)) + 
        facet_wrap(~county, ncol=5, labeller=labeller(county=countyLabels)) +
        labs(x=NULL, y="Cumulative indicence") + 
        theme(
            panel.grid.major.x=element_blank(),
            legend.title=element_blank(),
            axis.title.y=element_text(vjust=1, margin=margin(0, .5, 0, 0, "in")),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=90, hjust=1, vjust=1),
            axis.text.y=element_text(vjust=1),
            legend.position="top",
            strip.text.x=element_text(margin=margin(.2, 0, .1, 0, "in")),
            plot.margin=unit(c(0,1.5,.375,1.875),"in"),
            panel.border=element_rect(fill=NA, colour="grey70", size=rel(5)),
            panel.grid.major=element_line(size=rel(2.5)),
            panel.grid.minor=element_line(size=rel(1.25)),
            axis.ticks=element_line(colour="grey70", size=rel(2.5)),
        )

    dev.off()
@

<<coverageByRegionPoster, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/coverageByRegionRegimenPoster.tex", figuresDir), width=posterPlotWidth, height=posterPlotHeight * 1.25, pointsize=37)
    data = unprotectedByCounty %>% 
        filter(county %in% countiesForPlots, rounding==0, sliding==FALSE)

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        scale_y_continuous(labels=latexPercent, limits=c(0.9, 1), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        scale_color_manual(name="Prophylaxis regimen", labels=legendLabels(stratLabels[c(seq(1, 6), 10)], 15), values=c(rep(c(darkBrown, darkOrange), each=3), darkPurple)) + 
        scale_shape_manual(name="Prophylaxis regimen", labels=legendLabels(stratLabels[c(seq(1, 6), 10)], 15), values=c(16, 15, 17, 16, 15, 17, 16)) + 
        geom_errorbar(
            aes(ymin=frac.lower, ymax=frac.upper, x=county, color=strat, group=strat), 
            width=.3, 
            lwd=2,
            position=position_dodge(.5)
        ) + 
        geom_point(aes(y=frac.median, x=county, color=strat, shape=strat), position=position_dodge(.5), size=.75*5) +
        labs(y="Preventable fraction", color="Prophylaxis regimen") +
        guides(color=guide_legend(keywidth=3, keyheight=.5, ncol=3, title.hjust=1)) + 
        scale_x_discrete(labels=countyLabels) + 
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.title.y=element_text(vjust=1, margin=margin(0, .5, 0, 0, "in")),
            legend.position="top",
            axis.text.x=element_text(vjust=1, margin=margin(.125, 0, 0, 0, "in")),
#            strip.text.x=element_text(margin=margin(.2, 0, .1, 0, "in")),
            plot.margin=unit(c(0,1.5,.375,1.875),"in"),
            panel.border=element_rect(fill=NA, colour="grey70", size=rel(5)),
            panel.grid.major=element_line(size=rel(2.5)),
            panel.grid.minor=element_line(size=rel(1.25)),
            axis.ticks=element_line(colour="grey70", size=rel(2.5)),
        )
    dev.off()
@

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\edef\figuresDir{.texpadtmp/figures}

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamertemplate{headline}{
 \leavevmode
  \begin{columns}
   \begin{column}{\linewidth}
    \centering
    \usebeamercolor{title in headline}{\color{jblue}\Huge{\textbf{\inserttitle}}\\[0.5ex]}
    \usebeamercolor{author in headline}{\color{fg}\Large{\insertauthor}\\[.5ex]}
%    \usebeamercolor{institute in headline}{\color{fg}\large{\insertinstitute}\\[.5ex]}
    \vskip1cm
   \end{column}
   \vspace{1cm}
  \end{columns}
 \vspace{0.5in}
 \hspace{0.5in}\begin{beamercolorbox}[wd=46in,colsep=0.15cm]{cboxb}\end{beamercolorbox}
 \vspace{0.1in}
}

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=white} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{47in} % A0 width: 46.8in
\setlength{\paperheight}{35in} % A0 height: 33.1in
\setlength{\sepwid}{0pt} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.44\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{Assessment and optimization of RSV prophylaxis in Connecticut} % Poster title

\author{Ben Artin, YSPH/YSM 2018} % Author(s)

\institute{Advisors: Daniel Weinberger, PhD; Virginia Pitzer, ScD} % Institution(s)

%----------------------------------------------------------------------------------------

\usepackage[backend=bibtex,style=nejm,sorting=none,autocite=superscript,articledoi=true,url=false]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)

\addbibresource{../Thesis.bib} % The filename of the bibliography
%\renewcommand{\cite}[1]{\autocite{#1}}
\renewcommand{\cite}[1]{}

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

\begin{alertblock}{Objectives}

\begin{itemize}
\item Estimate variation in timing of respiratory syncytial virus (RSV) season in Connecticut by county and by year
\item Evaluate effectiveness of the RSV immunoprophylaxis regimen recommended by the American Academy of Pediatrics (AAP)
\item Formulate alternative prophylaxis regimens to account for regional and annual variation in timing of RSV season in Connecticut
\item Compare effectiveness of alternative regimens against the AAP recommendation
\end{itemize}

\end{alertblock}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Background}

Respiratory syncytial virus (RSV) causes seasonal respiratory infection with potentially serious complications in children,\cite{Hall:2009gi,Hall:2013jo,Zhou:2012ho,Boyce:2000gz} and leads to hospitalization rates as high as 50\% in high-risk infants.\cite{AmericanAcademyofPediatricsCommitteeonInfectiousDiseases:2014bj,Boyce:2000gz} Safe and effective immunoprophylaxis (palivizumab) is available, but costly.\cite{Hampp:2011ju,Andabaka:2013dr,Group:1998ih,Feltes:2003wm}

The American Academy of Pediatrics (AAP) recommends palivizumab only for infants at high risk of complications, and only during the RSV season.\cite{Anonymous:1998up,AmericanAcademyofPediatricsCommitteeonInfectiousDiseasesandCommitteeonFetusandNewborn:2003ug,AmericanAcademyofPediatricsCommitteeonInfectiousDiseases:2014bj} Although the current AAP guidelines acknowledge the existence of spatial and temporal variation in RSV incidence, they do not recommend spatial or temporal adjustments to immunoprophylaxis regimen outside of Florida.\cite{CentersforDiseaseControlandPreventionCDC:2011wt,CentersforDiseaseControlandPreventionCDC:2013wv,AmericanAcademyofPediatricsCommitteeonInfectiousDiseases:2014bj}

In this study, we investigated the value of using spatial and temporal variation in RSV incidence to adjust the RSV prophylaxis regimen in Connecticut.

\begin{center}
\vspace{2.35in}
\includegraphics[height=320pt]{EPH}
\hspace{40pt}
\includegraphics[height=320pt]{Medicine}
\end{center}

\end{block}

%------------------------------------------------

%\begin{figure}
%\includegraphics[width=0.8\linewidth]{placeholder.jpg}
%\caption{Figure caption}
%\end{figure}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

%\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column
%
%\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)
%
%\end{column} % End of column 2.1
%
%\begin{column}{\onecolwid}\vspace{-.6in} % The second column within column 2 (column 2.2)
%
%\begin{block}{Methods}
%
%\end{block}
%
%%----------------------------------------------------------------------------------------
%
%\end{column} % End of column 2.2
%
%\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width

\begin{block}{Methods}

\begin{enumerate}
\item Identified \Sexpr{nrow(dataset)} RSV-associated hospitalizations in Connecticut between \Sexpr{formatMonth(startDate)} and \Sexpr{formatMonth(endDate)}. 
\item Modeled RSV seasons using Monte Carlo sampling of log-linked generalized additive model with penalized cubic B-splines.
%\item We developed and published a custom R package, \texttt{outbreak-inference}, for estimation of outbreak characteristics.
\item Estimated the fraction of all RSV-associated hospitalizations that occur during the interval when RSV prophylaxis is recommended (``preventable fraction'') under the AAP guidelines.
\item Estimated RSV season onset and offset in each region across all years.
\item Constructed alternative prophylaxis regimens based on regional variation in RSV season timing.
\item Estimated preventable fraction of the regionally adjusted alternative regimens.
\end{enumerate}

\end{block}

\begin{block}{Results}

\begin{alertblock}{AAP-recommended prophylaxis lags behind RSV season}

\begin{figure}
\begin{center}
\input{\figuresDir/coverageByRegionAAPInsightPoster.tex}
\label{fig:coverageByRegionAAPInsight}
\caption{Regional variation in the lag between RSV seasons and the AAP-recommended prophylaxis regimen}
\end{center}
\end{figure}

\end{alertblock}

\begin{alertblock}{Spatial adjustments yield up to 1\% improvement}

\begin{figure}
\begin{center}
\input{\figuresDir/coverageByRegionRegimenPoster.tex}
\label{fig:coverageByRegionRegimen}
\caption{Comparison of six spatially adjusted RSV prophylaxis regimens to the AAP-recommended regimen}
\end{center}
\end{figure}

\end{alertblock}
\end{block}

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------

\end{column} % End of the second column

\begin{column}{\onecolwid} % The third column

%----------------------------------------------------------------------------------------
%	CONCLUSION
%----------------------------------------------------------------------------------------

\begin{block}{Conclusion}

Timing of RSV season in Connecticut varies by several weeks in different parts of the state. The AAP-recommended
prophylaxis regimen is misaligned with the RSV season in the counties with highest RSV incidence.

Timing of RSV immunoprophylaxis could be adjusted to achieve approximately a 1\% increase in RSV prevention statewide.

This gain requires using immunoprophylaxis schedule based on 1-week or 2-week intervals, rather 
than the month-based schedule currently used in AAP recommendations.

Year-to-year variability of RSV season timing in Connecticut is less significant than regional differences within the state.
Accounting for year-to-year variability in our model did not produce any additional improvements beyond the 1\% gained from accounting 
for regional variability.

\end{block}

%----------------------------------------------------------------------------------------
%	ADDITIONAL INFORMATION
%----------------------------------------------------------------------------------------

\begin{block}{Limitations}

Our model of AAP prophylaxis regimen assumes protection begins on November 15th, whereas prophylaxis start dates
are actually spread out throughout November.

We assumed that RSV illness after receiving prophylaxis (breakthrough illness) is negligible.

We do not have data for RSV infections among travelers from other states, and therefore our analysis underestimates RSV
incidence in regions with high levels of interstate travel, such as urban centers near state lines.

Our regional analysis does not account for inter-county travel, such as in employment hubs near county lines.

\end{block}

%----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

%\begin{block}{References}
%
%\small{\printbibliography}
%
%\end{block}

%----------------------------------------------------------------------------------------
%	CONTACT INFORMATION
%----------------------------------------------------------------------------------------

%\begin{center}
%\begin{tabular}{ccc}
%\includegraphics[width=0.4\linewidth]{logo.png} & \hfill & \includegraphics[width=0.4\linewidth]{logo.png}
%\end{tabular}
%\end{center}

\setbeamercolor{block alerted title}{fg=black,bg=black!10} % Change the alert block title colors
\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors

\vspace{2.1in}

\begin{alertblock}{Thesis advisors}

\begin{itemize}
\item Daniel Weinberger, PhD, YSPH
\item Virginia Pitzer, ScD, YSPH
\end{itemize}

\end{alertblock}

%----------------------------------------------------------------------------------------

\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}