### 2019-03-01 Paper/RSVOptimization.rnw
Preventable fraction plots in supplement

\documentclass[10pt,letter]{article}
\usepackage{float}
\floatplacement{figure}{H}

\usepackage[toc,page]{appendix}
\renewcommand{\appendixpagename}{Supplements}
\renewcommand{\appendixname}{Supplement}

\usepackage{printlen}
\usepackage{textcomp}
\usepackage{setspace}

\usepackage[hidelinks]{hyperref}
\usepackage[svgnames]{xcolor} % Must go before tikz and before structure.tex
\usepackage{tikz}
\usepackage{draftwatermark}
\SetWatermarkScale{1}
\SetWatermarkAngle{45}
\SetWatermarkFontSize{4ex}
\SetWatermarkHorCenter{12ex}
\SetWatermarkVerCenter{12ex}
\SetWatermarkLightness{.5}
\SetWatermarkText{DRAFT 1}

\input{structure.tex}
\usepackage{amssymb}
\usepackage{wasysym}

\usepackage[bf,medium,raggedright,compact]{titlesec}
\titleformat{\section}{\normalfont\large\bfseries\scshape}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\normalsize\bfseries\scshape}{}{0em}{}
\titleformat{\paragraph}{\normalfont\normalsize\bfseries}{}{0em}{}
\titlespacing*{\paragraph}{0pt}{3.25ex plus 1ex minus .2ex}{0em}

\usepackage{hanging}
\usepackage{graphicx}

\newlength\abstractindent
\setlength\abstractindent{1in}
\newcommand{\abstractsection}[1]{\hangpara{\abstractindent}{1}\makebox[1in][l]{\textbf{\textsc{#1}}}}
\newcommand{\abstractpara}{\hskip\abstractindent\hskip\baseparindent}

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

<<options, echo=FALSE>>=
    knitr::opts_chunk$set(echo=FALSE, message=FALSE, error=FALSE)
@

<<load>>=
knitr::read_chunk("../Analysis/RSV.R")
knitr::read_chunk("../Paper/Paper.R")
knitr::read_chunk("../Paper/Supplement.R")
@

<<libraries, cache=FALSE>>=
@

<<external, cache=TRUE>>=
@

<<paper, warning=FALSE>>=
@

\edef\figuresDir{\Sexpr{figuresDir}}
\title{Assessment and optimization of respiratory syncytial virus prophylaxis regimens in Connecticut, \Sexpr{formatYear(startDate)}-\Sexpr{formatYear(endDate)}}
\def\headertitle{Assessment and optimization of RSV prophylaxis regimens in Connecticut, \Sexpr{formatYear(startDate)}-\Sexpr{formatYear(endDate)}}

\author{
    \authorstyle{Ben Artin, MEng, MPH, MMSc\textsuperscript{1}}
    \newline\newline
    \authorstyle{Daniel Weinberger, PhD\textsuperscript{1}}\\
    \authorstyle{Virginia Pitzer, ScD\textsuperscript{1}}
    \newline\newline
    \textsuperscript{1}\institution{Yale University, School of Public Health}
}

\date{\today}

\pagestyle{empty}
\begin{document}
% \SweaveOpts{concordance=TRUE}

\newlength\baseparindent
\setlength\baseparindent{\parindent}
\begingroup
\setlength\parindent{0em}

\maketitle

\thispagestyle{firstpage}

\clearpage
\endgroup

\section*{\centerline{Abstract}}
\begin{hangparas}{\abstractindent}{1}
\begin{onehalfspacing}

\abstractsection{Background}Respiratory syncytial virus (RSV) causes seasonal respiratory infection with potentially serious complications in children, and leads to hospitalization rates of up to 50\% in high-risk infants. Safe and effective immunoprophylaxis (palivizumab) is available, but costly. The American Academy of Pediatrics (AAP) recommends palivizumab only for infants at high risk of complications, and only during the RSV season. Although the current AAP guidelines acknowledge the existence of spatial and temporal variation in RSV incidence, they do not recommend spatial or temporal adjustments to the immunoprophylaxis regimen outside of Florida. In this study, we investigate the value of using spatial and temporal variation in RSV incidence to adjust the RSV prophylaxis regimen in Connecticut.

\abstractsection{Methods}We obtained weekly ZIP-code-level counts of RSV-associated hospital admission in Connecticut between \Sexpr{formatMonth(startDate)} and \Sexpr{formatMonth(endDate)}. We fit smoothed curves to these admission counts and used these curves to estimate the fraction of all RSV cases in Connecticut occurring during the period of protection offered by RSV immunoprophylaxis (``preventable fraction'') under the AAP guidelines. We then used the same model to estimate protection conferred by immunoprophylaxis regimens with alternate start dates (but unchanged duration and net pharmaceutical cost).

\abstractsection{Results}Using preventable fraction to assess different immunoprophylaxis regimens, we found that regimens adjusted for county-level variation in timing of RSV seasons are superior to the current AAP-recommended regimen. We also considered the effect of rounding the timing of the first dose of prophylaxis to pragmatic calendrical boundaries (weekly, biweekly, and monthly), and found that benefits of alternative regimens over the AAP-recommended regimen persisted with biweekly rounding, but not with monthly rounding. Our best-performing pragmatic alternative to the AAP guidelines was based on the regional RSV season midpoint with biweekly rounding. In comparison to the AAP recommendation, whose preventable fraction is \Sexpr{regionStrategyCoverage("all", "aap", 0)} (\Sexpr{regionStrategyCoverageCI("all", "aap", 0, level)}), that alternative yielded improvement to \Sexpr{regionStrategyCoverage("all", "countyMiddle", 2)} (\Sexpr{regionStrategyCoverageCI("all", "countyMiddle", 2, level)}). We also found that alternative regimens adjusted for annual variation in the RSV season are non-superior to spatially adjusted regimens.

\abstractsection{Conclusion}Overall, we recommend county-level spatial analysis of RSV incidence as the starting point for RSV immunoprophylaxis optimization in Connecticut. However, the potential reduction in RSV hospitalizations should be weighed against the potential increase in implementation cost.

\abstractpara{}

\end{onehalfspacing}
\end{hangparas}

\clearpage

\pagestyle{fancy}

%\listoftables
%\clearpage

\listoffigures
\clearpage

\section{Introduction}

\subsection{Respiratory syncytial virus}

Respiratory syncytial virus (RSV) causes upper respiratory infections in humans of all ages. In adults, RSV infection is typically asymptomatic or gives rise to mild respiratory illness. However, RSV infection often incites more serious respiratory illness in children, with hospitalization rates as high as 4.4\% in infants with no comorbidities, making RSV a leading cause of hospitalization among infants.\cite{Hall:2009gi,Hall:2013jo,Zhou:2012ho,Boyce:2000gz}

For infants, risk factors of serious illness and hospitalization due to RSV infection include prematurity, chronic lung disease of prematurity, congenital heart disease, anatomic pulmonary abnormalities, neuromuscular disorders, trisomy 21, and immunocompromised status.\cite{AmericanAcademyofPediatricsCommitteeonInfectiousDiseases:2014bj} The hospitalization rate due to RSV among high-risk infants is anywhere from two to ten times higher than among infants with no comorbidities.\cite{Boyce:2000gz}

Since 1998, a monoclonal antibody (palivizumab), has been approved in the US for prevention of RSV infection in high-risk infants less than 2 years old. Palivizumab is administered via intramuscular injection and requires five monthly doses, priced at \$1500-\$3000 per dose.\cite{Hampp:2011ju,Andabaka:2013dr} Palivizumab has been found to be effective in reducing hospitalizations due to RSV; in double-blinded trials in high-risk infants, it reduced the hospitalization rate by 45-55\% compared to placebo.\cite{Group:1998ih,Feltes:2003wm} It has also been found to be safe and well-tolerated in clinical trials.\cite{Group:1998ih} Owing to the high cost, the 2014 revision of the American Academy of Pediatrics (AAP) RSV prophylaxis guidelines, recommends palivizumab only for high-risk infants, and only during the RSV season.\cite{Anonymous:1998up,AmericanAcademyofPediatricsCommitteeonInfectiousDiseasesandCommitteeonFetusandNewborn:2003ug,AmericanAcademyofPediatricsCommitteeonInfectiousDiseases:2014bj}

RSV seasonality and spatial variability have been established throughout the United States, based on National Respiratory and Enteric Virus Surveillance System data. In all states except for Florida, RSV shows an annual or biennial cycle with a stable seasonal onset between mid-November and early January and a stable seasonal offset between mid-March and late April.\cite{CentersforDiseaseControlandPreventionCDC:2011wt,CentersforDiseaseControlandPreventionCDC:2013wv} RSV season in Florida begins approximately two months earlier than in other states. \cite{CentersforDiseaseControlandPreventionCDC:2013wv}

The AAP acknowledges the potential significance of spatial and temporal variation in RSV incidence, but --- outside of Florida --- does not recommend local variance from nationwide prophylaxis guidelines.\cite{AmericanAcademyofPediatricsCommitteeonInfectiousDiseases:2014bj}

Previous analysis of RSV-associated hospital admissions in Connecticu has shown regional variability in time of peak RSV incidence within Connecticut.\cite{Noveroske:y4fi3188} No prior research is available regarding year-to-year variation in timing of RSV season in Connecticut.

\subsection{Study aims}

Motivated by the AAP's recognition of the significance of spatial variation in formulating prophylaxis guidelines, and the known spatial variability of RSV incidence in Connecticut, we aim to determine whether spatial and temporal variability can be used to optimize RSV prophylaxis in Connecticut by

\begin{itemize}
	\item To what extent does regional variability of RSV incidence in Connecticut impact the effectiveness of the AAP-recommended prophylaxis regimen?
	\item Can we use spatial analysis of RSV to propose alternative regimens that would increase the benefits of prophylaxis in comparison to the AAP-recommended regimen, without increasing the total cost of the regimen?
	\item Is there year-to-year variation in timing of RSV season in Connecticut?
	\item Can we take advantage of any year-to-year variation to improve upon AAP-recommended prophylaxis, without increasing treatment cost?
\end{itemize}

\section{Methods}

\subsection{Case definition}

In Connecticut, the Department of Public Health (CT DPH), via its Office of Health Care Access (OHCA), maintains a database of all hospitalizations in every non-federal short-stay acute-care general hospital in the state. We defined an RSV-associated hospitalization in Connecticut to be a hospital stay for which:

\begin{itemize}
    \item the postal code of residence listed the in patient's medical record was in Connecticut, and
    \item the hospital stay was associated with ICD-9-CM diagnosis code \texttt{079.6} (RSV), \texttt{466.11} (Acute bronchiolitis due to RSV), or \texttt{480.1} (Pneumonia due to RSV).
\end{itemize}

Using the OHCA weekly hospitalization statistics, we identified \Sexpr{prettyNum(nrow(dataset), big.mark=",")} RSV-associated hospitalizations in Connecticut between \Sexpr{formatMonth(startDate)} and \Sexpr{formatMonth(endDate)} (inclusive).

The study was approved by the Human Investigation Committees at Yale University and the CT DPH. The authors assume full responsibility for analyses and interpretation of the data obtained from CT DPH.

% TODO patient age

\subsection{Data normalization}

We designated epidemiological years as beginning in the first full week of July, in order to accommodate the timing of RSV peaks (which span two calendar years) and the limitations of hospitalization records (which OHCA aggregated into Monday-Sunday calendar weeks). Due to changes in diagnostic coding, we excluded data prior to July 1996; we similarly excluded data after June 2013 due to incompleteness.

\subsection{Choice of model}

Our goal was to construct a model of annual RSV incidence that allowed us to obtain interval estimates of complex characteristics of RSV seasons, such as season onset, season offset, and preventable fraction of a prophylaxis regimen. To that end, we fit smooth curves through annual RSV hospitalization data, and used resampling to obtain interval estimates. (See Supplement \ref{supplement:splines}.)

\subsection{Spatial aggregation of RSV incidence}

<<map, results='hide'>>=
    tikz(sprintf("%s/map.tex", figuresDir), width=inlinePlotWidth, height=inlinePlotHeight, pointsize=10)
    US.counties = readOGR(dsn="../Maps/cb_2016_us_county_5m",layer="cb_2016_us_county_5m")
    US.counties = US.counties[US.counties$STATEFP == "09",]
    county.data = US.counties@data %>%
        mutate(id=rownames(.)) %>%
        data.table() %>%
        mutate(highRSV=ifelse(NAME %in% lowIncidenceCounties, FALSE, TRUE))

    map.df = US.counties %>% 
        fortify() %>% 
        data.table() %>%
        inner_join(county.data %>% select(highRSV, id), by="id")

    county.data = map.df %>%
        group_by(id) %>%
        summarize(lat=mean(range(lat)), long=mean(range(long))) %>%
        ungroup() %>%
        inner_join(county.data, by="id") %>%
        rename(name=NAME)

    ggplot(map.df, aes(x=long, y=lat)) +
        theme_light(base_size=plotTextBaseSize) +
        scale_fill_manual(labels=legendLabels(c("$<$50", "$>$100")), values=c(lightBlue, darkBlue)) +
        geom_polygon(aes(fill=highRSV, group=group), color="white", size=.5) +
        geom_text(data=county.data, aes(label=name), size=2.2) +
        coord_map() + 
        labs(fill="Average RSV-associated hospitalizations per year") +
        theme(
            rect=element_blank(),
            line=element_blank(),
            axis.text=element_blank(),
            axis.title=element_blank(),
            legend.position="bottom"
        )
    dev.off()
@

For our spatial analysis, we aggregated RSV incidence data between \Sexpr{formatYear(startDate)} and \Sexpr{formatYear(endDate)} into five regions:

\begin{itemize}
    \item One region for the entire state
    \item One region for each of the counties with high RSV case counts: Hartford County, Fairfield County, and New Haven County (Figure \ref{fig:map}).
    \item One discontiguous region comprised of all remaining counties in Connecticut: \Sexpr{gsub("(.*), (.*)", "\\1, and \\2", paste(lowIncidenceCounties, collapse=", "))} (``low-RSV counties'').
\end{itemize}

We chose county-level aggregation with an eye towards using this data to inform clinical practice guidelines, which are typically based on administrative regions for reasons of administrative and implementation simplicity.

We chose to aggregate low-RSV counties into a single region because our analysis in those counties individually yielded results with low statistical power.

\begin{figure}
\begin{center}
\input{\figuresDir/map.tex}
\caption{Subdivision of Connecticut into high-RSV and low-RSV counties}
\label{fig:map}
\end{center}
\end{figure}

\subsection{Evaluating the AAP prophylaxis recommendation}

We began by assessing the current AAP RSV prophylaxis guidelines. These guidelines recommend RSV prophylaxis with 5 monthly doses of palivizumab, starting in November, for all high-risk infants in their first year of life; in the second year of life, palivizumab is only recommended for some subgroups. For simplicity, we assumed that protection by palivizumab begins on November 15th and lasts 24 weeks.

We defined the fraction of all RSV cases that is preventable by a prophylaxis regimen (``preventable fraction'') to be the ratio of

\begin{itemize}
\item the number of RSV-associated hospitalizations that occur while the prophylaxis regimen offers protection to those who receive prophylaxis as scheduled, and
\item the total number of RSV-associated hospitalizations.
\end{itemize}

%TODO citation needed

Given that high-risk infants are a small subgroup of the general population, we assumed that our model of RSV hospitalizations among all infants is an unbiased approximation of RSV hospitalizations among high-risk infants, and that, therefore, the preventable fraction of a prophylactic regimen is an unbiased estimate of the fraction of high-risk infants who benefit from prophylaxis.

\subsection{Estimating all-years characteristics of RSV seasons}

We then turned our attention to spatial variation of seasonal RSV patterns between \Sexpr{formatMonth(startDate)} and \Sexpr{formatMonth(endDate)}. In particular, we defined season onset as the time when cumulative incidence rises beyond \Sexpr{seasonThreshold * 100}\% of the total cumulative incidence, and season offset as the time when cumulative incidence rises beyond \Sexpr{(1 - seasonThreshold) * 100}\% of the total cumulative incidence. We aggregated our data by surveillance week across all surveillance years, and estimated RSV season onset and offset for each region.

The choice of \Sexpr{seasonThreshold * 100}\% and \Sexpr{(1 - seasonThreshold) * 100}\% incidence thresholds was based on our intent to use estimates of RSV season onset and offset to put forth alternative prophylaxis regimens (as described below), while maintaining the 5-month dosing schedule (and therefore not increasing medication cost). Incidence cutoffs other than \Sexpr{seasonThreshold * 100}\% and \Sexpr{(1 - seasonThreshold) * 100}\% were briefly investigated, but quickly ruled out because they produced prophylaxis regimens starting long prior to RSV season onset or ending long after RSV season offset, and therefore had very low preventable fractions.

\subsection{Evaluating spatially adjusted prophylaxis regimens}

Following this analysis, we estimated the regional and statewide preventable fraction for six spatially adjusted prophylaxis regimens:

\begin{itemize}
    \item \textbf{By statewide all-years onset}: prophylaxis administration begins at statewide all-years median RSV season onset. 
    \item \textbf{By statewide all-years midpoint}: prophylaxis administration begins 12 weeks before the average of the (statewide all-years median) season onset and offset. 
    \item \textbf{By statewide all-years offset}: prophylaxis administration begins 24 weeks before the (statewide all-years median) RSV season offset. 
    \item \textbf{By regional all-years onset}: prophylaxis administration begins at regional\footnote{county-wide, or --- in the case of low-RSV counties --- low-RSV-region-wide} all-years median RSV season onset.
    \item \textbf{By regional all-years midpoint}: prophylaxis administration begins 12 weeks before the average of the (regional all-years median) RSV season onset and offset. 
    \item \textbf{By regional all-years offset}: prophylaxis administration begins 24 weeks before the (regional all-years median) RSV season offset. 
\end{itemize}

In all our alternative regimens, prophylaxis consists of five monthly doses of palivizumab, and protection is assumed to last 24 weeks from the first dose.

\subsection{Accommodating implementation complexity}

Keeping in mind implementation constraints of clinical practice guidelines, we performed three additional variants of our spatial analysis:

\begin{itemize}
    \item \textbf{Weekly rounding}: prophylaxis start and end date rounded to the nearest calendar week,
    \item \textbf{Biweekly rounding}: prophylaxis start and end date rounded to the nearest two-week calendar period, with periods chosen to approximately align with November 15th, and
    \item \textbf{Monthly rounding}: prophylaxis start and end date rounded to the nearest four-week calendar periods, with periods chosen to approximately align with November 15th.
\end{itemize}

In each variant, we again estimated the preventable fraction of all six spatially adjusted regimens.

\subsection{Temporal aggregation of RSV incidence}

Our analysis thus far assumed that RSV seasons are similar year-to-year. To analyze the effects on prophylaxis regimens of year-to-year variation between RSV seasons, we chose to aggregate our data into (overlapping) periods of three consecutive years --- in order to maintain a higher level of statistical power, as well as to limit the impact of short-term variation on our results.

TODO This gave us \Sexpr{year(endDate) - year(startDate) -1} three-year periods, with the \Sexpr{formatYear(startDate + 2*365)} period covering surveillance years \Sexpr{formatYear(startDate)} through \Sexpr{formatYear(startDate + 2*365)}, and the \Sexpr{formatYear(endDate)} period covering \Sexpr{formatYear(endDate - 2*365)} through \Sexpr{formatYear(endDate)}.

We began our temporal analysis by estimating onset and offset for each three-year period, using the same penalized spline GAM as for the all-years analysis, producing --- as before --- statewide and regional estimates of \Sexpr{seasonThreshold * 100}\% onset and \Sexpr{(1 - seasonThreshold) * 100}\% offset.

To evaluate year-to-year variation in RSV season onset and offset we modeled median RSV three-year onset and offset against time using simple linear regression.

We did not analyze any other long-term patterns, such as cyclicity.

\subsection{Evaluating alternative recent-years prophylaxis regimens}

Finally, to assess the impact of any long-term trends in RSV season timing on prophylaxis, we evaluated three additional prophylaxis regimens. Here, each prophylaxis regimen is adjusted annually based on RSV data from the preceding three-year period:

\begin{itemize}
    \item \textbf{By statewide recent-years onset}: prophylaxis administration in a given year begins at statewide median onset from the preceding three years. 
    \item \textbf{By statewide recent-years midpoint}: prophylaxis administration begins 12 weeks before the average of the (statewide median) onset and offset from the preceding three years. 
    \item \textbf{By statewide recent-years offset}: prophylaxis administration in a given year begins 24 weeks before the (statewide median) offset from the preceding three years.
\end{itemize}

\begin{figure}[b]
\begin{center}
\input{\figuresDir/coverageByRegionAAPInsight.tex}
\caption[Regional variation in the relative timing of the RSV season in Connecticut and the AAP-recommended prophylaxis regimen]{Regional variation in the relative timing of the RSV season in Connecticut and the AAP-recommended prophylaxis regimen, showing the gap between season onset and prophylaxis start date, notable particularly in Fairfield county}
\label{fig:coverageByRegionAAPInsight}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\input{\figuresDir/seasonByRegion.tex}
\caption{Regional variation in RSV season onset and offset in Connecticut}
\label{fig:seasonByRegion}
\end{center}
\end{figure}

We then compared those regimens to the one recommended by the AAP and to the six spatially adjusted alternatives (with no calendrical rounding) discussed above.

As before, these alternative regimens involved five monthly doses of palivizumab giving 24 weeks of protection.

% \uselengthunit{in}\printlength{\columnwidth}
% column width = 3.20929 in

\section{Results}

\paragraph{RSV season timing varies regionally}

<<seasonByRegion, results='hide'>>=
    tikz(sprintf("%s/seasonByRegion.tex", figuresDir), width=inlinePlotWidth, height=inlinePlotHeight, pointsize=10)
    data = thresholdsByCounty %>% 
        filter(county %in% countiesForPlots) %>%
        melt(id=c("county")) %>%
        cSplit(c("variable"), c(".")) %>%
        rename(variable=variable_1, pctile=variable_2) %>%
        mutate(variable=factor(variable, levels=c("offset", "onset"))) %>%
        cast(county + variable ~ pctile)

    onsetOffsetRange = 6 # weeks
    onsetOffsetAdjust = data.frame(variable=c("onset", "offset"), nudge=c(-1, -2))

    limits = data %>% 
        select(variable, median) %>%
        group_by(variable) %>%
        summarize(mid=round(mean(range(median)))) %>%
        inner_join(onsetOffsetAdjust, by="variable") %>%
        mutate(mid=mid + nudge, min=mid - onsetOffsetRange / 2, max=mid + onsetOffsetRange / 2)

    monthBarX = 0.5 + (0.5 - 5.5) * 0.04
    monthGrob = function(month) {
        textGrob(epiWeekLabels[month], just="right", gp=gpar(fontsize="6.4", col="grey30"))
    }

    plot = ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=county), width=.1) + 
        geom_point(aes(y=median, x=county), size=.75) + 
        geom_blank(data=limits, aes(y=min)) + 
        geom_blank(data=limits, aes(y=max)) + 
        labs(aes(x=variable)) +
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        scale_x_discrete(labels=countyLabels) + 
        scale_y_continuous(minor_breaks=seq(1, 52), breaks=epiWeekBreaks, labels=rep("\\qquad\\qquad", length(epiWeekBreaks)), sec.axis=sec_axis(trans=~., breaks=seq(1, 52))) + 
        labs(y=NULL) + 
        annotation_custom(monthGrob(4), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[4], ymax=monthBoundaries$mid[4]) + 
        annotation_custom(monthGrob(5), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[5], ymax=monthBoundaries$mid[5]) + 
        annotation_custom(monthGrob(9), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[9], ymax=monthBoundaries$mid[9]) + 
        annotation_custom(monthGrob(10), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[10], ymax=monthBoundaries$mid[10]) + 
        theme(
            panel.grid.major.y=element_blank(),
            panel.grid.minor.y=element_line(size=rel(2)),
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=30, hjust=1)
        )

	plotGrob = ggplotGrob(plot)
	plotGrob$layout$clip[grepl("panel", plotGrob$layout$name)] <- "off"
	grid.draw(plotGrob)

    dev.off()
@

Compared to the statewide median onset of RSV season at \Sexpr{regionOnset("all")} weeks (\Sexpr{regionOnsetCI("all", level)} weeks) and offset at \Sexpr{regionOffset("all")} weeks (\Sexpr{regionOffsetCI("all", level)} weeks), RSV season in Fairfield county has earlier median onset at \Sexpr{regionOnset("Fairfield")} weeks (\Sexpr{regionOnsetCI("Fairfield", level)} weeks) and later median offset at \Sexpr{regionOffset("Fairfield")} weeks (\Sexpr{regionOffsetCI("Fairfield", level)} weeks) (Figure \ref{fig:seasonByRegion}).

\begin{figure}
\begin{center}
\input{\figuresDir/coverageByRegionAAP.tex}
\caption[Regional variation in the fraction of RSV hospitalizations occurring while the AAP-recommended RSV prophylaxis is active]{Regional variation in the fraction of RSV hospitalizations occurring while protection by palivizumab --- administered per the AAP guidelines ---  is active}
\label{fig:coverageByRegionAAP}
\end{center}
\end{figure}

TODO Meanwhile, in comparison with the statewide RSV season onset and offset, the low-RSV region has RSV season with later onset at \Sexpr{regionOnset("lowIncidence")} weeks (\Sexpr{regionOnsetCI("all", level)} weeks); season offset shows no change from the statewide median (Figure \ref{fig:seasonByRegion}).

\paragraph{The AAP guidelines perform best in low-RSV counties and worst in Fairfield county}

<<coverageByRegionAAP, results='hide'>>=
    tikz(sprintf("%s/coverageByRegionAAP.tex", figuresDir), width=inlinePlotWidth, height=inlinePlotHeight, pointsize=10)

    data = unprotectedByCounty %>% 
        filter(strat=="aap", rounding==0) %>% 
        filter(county %in% countiesForPlots)

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymax=frac.upper, ymin=frac.lower, x=county), width=.1) + 
        geom_point(aes(y=frac.median, x=county), size=.75) + 
        scale_y_continuous(labels=latexPercent, limits=c(0.9, 1), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        scale_x_discrete(labels=countyLabels) + 
        labs(y="Preventable fraction") +
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=30, hjust=1)
        )
    dev.off()
@

<<coverageByRegionAAPInsight, results='hide'>>=
    tikz(sprintf("%s/coverageByRegionAAPInsight.tex", figuresDir), width=pagePlotWidth, height=pagePlotHeight, pointsize=10)

    strat = ppxFixedStrategies("all")[["aap"]]

    data = predByCounty %>%
	    filter(county %in% countiesForPlots)

    thresholds = data %>%
    	mutate(ppx=strat(time)) %>%
    	filter(ppx > 0) %>%
    	group_by(county) %>%
    	do(
    		data.frame(
    			ppx.start=min(.$time),
    			ppx.end=max(.$time),
    			unprotected.start=min(.$rsv.cum.fit),
    			unprotected.end=max(.$rsv.cum.fit)
    		)
    	) %>%
    	ungroup() %>%
    	as.data.frame() %>%
    	inner_join(thresholdsByCounty, by="county")

    barYmin = 0.025
    barYmax = 0.075

    spanExtents = data.frame(
        variable=c("ppx", "season"), 
        barYmin=c(1 + barYmin, -barYmax), 
        barYmax=c(1 + barYmax, -barYmin), 
        lineYmin=c(0, -barYmin), 
        lineYmax=c(1 + barYmin, 1)
    )

    spans = thresholds %>%
        select(county, matches("ppx."), matches(".median")) %>%
        rename(season.start=onset.median, season.end=offset.median) %>%
        melt(id.vars="county") %>%
        cSplit("variable", ".") %>%
        rename(variable=variable_1, part=variable_2) %>%
        cast(county + variable ~ part) %>%
        right_join(spanExtents, by="variable")

    spanLabels = legendLabels(c("AAP-recommended prophylaxis", "RSV season"), leading=1)

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_ribbon(aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper, x=time), fill=grey(.25)) +
    	geom_segment(aes(x=start, y=lineYmin, xend=start, yend=lineYmax, color=variable), data=spans, size=.375) + 
    	geom_segment(aes(x=end, y=lineYmin, xend=end, yend=lineYmax, color=variable), data=spans, size=.375) + 
    	geom_rect(aes(xmin=start, ymin=barYmin, xmax=end, ymax=barYmax, fill=variable, color=variable), data=spans, size=.375) + 
    	geom_segment(aes(x=-Inf, y=unprotected.start, xend=ppx.start, yend=unprotected.start), linetype="11", data=thresholds, size=.375) + 
        geom_segment(aes(x=-Inf, y=unprotected.end, xend=ppx.end, yend=unprotected.end), linetype="11", data=thresholds, size=.375) + 
        geom_point(aes(x=min(monthBoundaries$min), y=unprotected.end), data=thresholds, shape=18) + 
        geom_point(aes(x=min(monthBoundaries$min), y=unprotected.start), data=thresholds, shape=18) + 
        scale_y_continuous(labels=latexPercent, breaks=seq(0, 1, by=.2), limits=c(-0.1, 1.1), expand=c(0, 0)) +
        scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
        scale_color_manual(labels=spanLabels, values=c(darkPurple, darkRed)) +
        scale_fill_manual(labels=spanLabels, values=c(darkPurple, darkRed)) +
        guides(color=guide_legend(keywidth=3, keyheight=.4)) + 
        facet_wrap(~county, ncol=5, labeller=labeller(county=countyLabels)) +
        labs(x=NULL, y="Cumulative indicence fraction") + 
        theme(
            panel.grid.major.x=element_blank(),
            legend.title=element_blank(),
            axis.title.y=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
            legend.position="bottom"
        )

    dev.off()
@

Statewide, our model shows that \Sexpr{regionStrategyCoverage("all", "aap", 0)} of RSV cases (\Sexpr{regionStrategyCoverageCI("all", "aap", 0, level)}) occur while palivizumab, administered per the AAP guidelines, offers protection.

\begin{figure}
\begin{center}
\input{\figuresDir/coverageByRegionRegimen.tex}
\caption[Comparison of six spatially adjusted RSV prophylaxis regimens to the AAP-recommended regimen]{Comparison of six spatially adjusted RSV prophylaxis regimens to the AAP-recommended regimen, showing statewide superiority and regional non-inferiority of spatially adjusted regimens. Preventable fraction: the fraction of all RSV-associated hospitalizations occurring while the prophylaxis regimen provides protection}
\label{fig:coverageByRegionRegimen}
\end{center}
\end{figure}

The AAP guidelines perform better in the low-RSV counties, where \Sexpr{regionStrategyCoverage("lowIncidence", "aap", 0)} cases (\Sexpr{regionStrategyCoverageCI("lowIncidence", "aap", 0, level)}) occur during the prophylaxis interval, but worse in Fairfield county, where \Sexpr{regionStrategyCoverage("Fairfield", "aap", 0)} cases (\Sexpr{regionStrategyCoverageCI("Fairfield", "aap", 0, level)}) occur during this time (Figure \ref{fig:coverageByRegionAAP}).

This is caused by temporal misalignment between the RSV season (delineated by median \Sexpr{seasonThreshold * 100}\%  onset and \Sexpr{(1 - seasonThreshold) * 100}\% offset in each region) and the timing of the AAP-recommended prophylaxis regimen (Figure \ref{fig:coverageByRegionAAPInsight}). Fairfield county, owing to its early season onset, sees disproportionately more cases before prophylaxis administration begins; on the other hand, the low-RSV counties, owing to their late season onset and relatively shorter RSV seasons, see almost all of their RSV cases within the prophylaxis window.

\paragraph{Spatially adjusted regimens are superior to the AAP recommendations}

<<coverageByRegion, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/coverageByRegionRegimen.tex", figuresDir), width=pagePlotWidth, height=pagePlotHeight * 1.5, pointsize=10)
    data = unprotectedByCounty %>% 
        filter(county %in% countiesForPlots, rounding==0, sliding==FALSE)

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        scale_y_continuous(labels=latexPercent, limits=c(0.9, 1), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        scale_color_manual(name="Prophylaxis regimen", labels=legendLabels(stratLabels[c(seq(1, 6), 10)]), values=c(rep(c(darkBrown, darkOrange), each=3), darkPurple)) + 
        scale_shape_manual(name="Prophylaxis regimen", labels=legendLabels(stratLabels[c(seq(1, 6), 10)]), values=c(16, 15, 17, 16, 15, 17, 16)) + 
        geom_errorbar(
            aes(ymin=frac.lower, ymax=frac.upper, x=county, color=strat, group=strat), 
            width=.3, 
            position=position_dodge(.5)
        ) + 
        geom_point(aes(y=frac.median, x=county, color=strat, shape=strat), position=position_dodge(.5), size=.75) +
        labs(y="Preventable fraction", color="Prophylaxis regimen") +
        guides(color=guide_legend(keywidth=3, keyheight=1, ncol=3)) + 
        scale_x_discrete(labels=countyLabels) + 
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            legend.position="bottom"
        )
    dev.off()
@

Based on all-years data, the six spatially adjusted prophylaxis regimens (with no rounding of prophylaxis dates), when compared to the AAP guidelines, yield non-inferior results in all regions, and superior results statewide (Figure \ref{fig:coverageByRegionRegimen}).

Statewide, the AAP guidelines offer protection to \Sexpr{regionStrategyCoverage("all", "aap", 0)} cases (\Sexpr{regionStrategyCoverageCI("all", "aap", 0, level)}). The alternative regimen based on all-years statewide onset yields the smallest increase, to \Sexpr{regionStrategyCoverage("all", "stateOnset", 0)} (\Sexpr{regionStrategyCoverageCI("all", "stateOnset", 0, level)}); other alternative regimens yield similar results.

Most of the increase comes from Hartford county, where the AAP guidelines protect \Sexpr{regionStrategyCoverage("Hartford", "aap", 0)} of cases (\Sexpr{regionStrategyCoverageCI("Hartford", "aap", 0, level)}), whereas the alternative regimen based on all-years regional offset generates the smallest increase, to \Sexpr{regionStrategyCoverage("Hartford", "countyOffset", 0)} (\Sexpr{regionStrategyCoverageCI("Hartford", "countyOnset", 0, level)}); other alternative regimens are comparable.

In every region, all alternative regimens based on all-years analysis are non-inferior to the AAP guidelines and non-superior to each other.

\paragraph{Calendrical rounding rapidly erases gains over the AAP-recommended regimen}

<<calendarForRounding, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/calendarForRoundingRegimen.tex", figuresDir), width=pagePlotWidth, height=pagePlotHeight * 1.5, pointsize=10)

    roundingStrat = unprotectedByCounty %>% 
        filter(county %in% countiesForPlots, sliding==FALSE) %>%
        select(strat, rounding) %>%
        unique()

    dataAll = predByCounty %>%
        filter(county %in% countiesForPlots) %>%
        group_by(county, time) %>%
        do(
            roundingStrat
        ) %>%
        ungroup() %>%
        group_by(county, strat, rounding) %>%
        do(
            (function(df) {
                county = unique(df$county)
                rounding = as.numeric(unique(df$rounding))
                strat = as.character(unique(df$strat))
                strat = ppxFixedStrategies(county, rounding)[[strat]]
                stratActive = strat(df$time)
                df %>%
                    mutate(stratActive=stratActive)
            })(.)
        ) %>%
        ungroup() %>%
        filter(stratActive > 0) %>%
        group_by(county, strat, rounding) %>%
        summarize(start=min(time), end=max(time)) %>%
        ungroup()

    aapStartEnd = dataAll %>%
        filter(strat == "aap") %>%
        select(start, end) %>%
        unique()

    aap = dataAll %>%
        filter(strat != "aap") %>%
        group_by(strat, rounding) %>%
        do(
            aapStartEnd
        ) %>%
        ungroup() %>%
        mutate(y=ifelse(grepl("state", strat), "all", "lowIncidence")) %>%
        mutate(y=factor(y, levels(dataAll$county)))

    data = dataAll %>% 
        filter(strat != "aap") %>%
        filter(xor(grepl("county", strat), county == "all"))

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
    	geom_rect(aes(xmin=start, xmax=end, ymin=as.numeric(county)-0.25, ymax=as.numeric(county)+0.25), size=.375) + 
    	geom_rect(data=aap, aes(xmin=start, xmax=end, ymin=as.numeric(y)+0.75, ymax=as.numeric(y)+1.25), size=.375) + 
        scale_y_continuous(breaks=as.numeric(countiesForPlots), labels=countyLabels[as.character(countiesForPlots)]) + 
        scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
#        scale_color_manual(labels=spanLabels, values=twoToneAlt) +
#        scale_fill_manual(labels=spanLabels, values=twoToneAlt)# +
#        guides(color=guide_legend(keywidth=3, keyheight=.4)) + 
        facet_wrap(~ rounding + strat, nrow=8, labeller=labeller(rounding=roundingLabels, strat=stratLabels), scales="free_y") #+
#        labs(x=NULL, y="Cumulative indicence fraction") + 
#        theme(
#            panel.grid.major.x=element_blank(),
#            legend.title=element_blank(),
#            axis.title.y=element_blank(),
#            axis.ticks.x=element_blank(),
#            axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
#            legend.position="bottom"
#        )
    dev.off()
@

<<coverageByRounding, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/coverageByRoundingRegimen.tex", figuresDir), width=pagePlotWidth, height=pagePlotHeight * 1.5, pointsize=10)

    # We are going to de-emphasize some data points, because sometimes different strategies under different rounding end up producing the same prophylaxis schedule, and there is no point in showing all of them equally on the plot. To do this, we consider the left-to-right order in which strategies will appear in the plot, and then de-emphasize any strategy whose schedule is the same as some strategy that appears on its left
    data = unprotectedByCounty %>% 
        filter(county == "all", sliding==FALSE) %>%
        arrange(rounding, strat) %>%
        mutate(strat=factor(strat, levels=c(levels(strat)[1], "_1", "_2", levels(strat)[2:7])))

    data$startIdx = match(data$desc, data$desc)
    data$isRepeat = data$startIdx != seq(1, length(data$startIdx))
    data$isRepeat = data$isRepeat & !data$strat == "aap" 
    data$strat.isRepeat = sprintf("%s.%s", data$strat, data$isRepeat)
    idx = order(rep(seq(1, length(levels(data$strat))), 2))
    strat.isRepeat.levels = (c(
      sprintf("%s.%s", levels(data$strat), FALSE), 
      sprintf("%s.%s", levels(data$strat), TRUE)
    ))[idx]
    data$strat.isRepeat = factor(data$strat.isRepeat, strat.isRepeat.levels)

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        scale_y_continuous(labels=latexPercent, limits=c(0.90, 1.00), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        scale_alpha_manual(guide=FALSE, breaks=c(FALSE, TRUE), values=c(1, 0.5)) + 
        scale_color_manual(
          name="Prophylaxis regimen", 
          breaks=factor(sprintf("%s.%s", levels(data$strat), FALSE), levels=strat.isRepeat.levels),
          labels=legendLabels(c(as.vector(stratLabels[10]), "_", "__", as.vector(stratLabels[1:6]))),
          values=c(
            darkPurple, darkPurple, 
            NA, NA, NA, NA, 
            rep(c(darkBrown, darkOrange), each=6)
          ), 
          drop=FALSE
        ) + 
        scale_shape_manual(
          name="Prophylaxis regimen", 
          breaks=factor(sprintf("%s.%s", levels(data$strat), FALSE), levels=strat.isRepeat.levels),
          labels=legendLabels(c(as.vector(stratLabels[10]), "_", "__", as.vector(stratLabels[1:6]))),
          values=c(
            16, 16, 
            16, 16, 16, 16, 
            16, 21, 15, 22, 17, 24, 
            16, 21, 15, 22, 17, 24
          ), 
          drop=FALSE
        ) + 
        scale_x_discrete(labels=roundingLabels) + 
        geom_errorbar(
            aes(ymin=frac.lower, ymax=frac.upper, x=rounding, color=strat.isRepeat, group=strat.isRepeat), 
            width=.25, 
            position=position_dodge(.5)
        ) + 
        geom_point(aes(y=frac.median, x=rounding, color=strat.isRepeat, shape=strat.isRepeat), position=position_dodge(.5), size=0.75, fill="white") +
        labs(x="Calendrical rounding", y="Preventable fraction") +
        guides(color=guide_legend(keywidth=3, keyheight=1, ncol=3)) + 
        theme(
            legend.position="bottom"
        )
    dev.off()
@

After we applied weekly, biweekly, and monthly rounding to start and end dates of our six alternative immunoprophylaxis regimens, we found that longer rounding intervals diminished the benefit of spatial analysis (Figure \ref{fig:coverageByRoundingRegimen}).

With weekly rounding, our regimens were non-inferior to the AAP recommendation, and produced an increase in preventable fraction from AAP's \Sexpr{regionStrategyCoverage("all", "aap", 0)} (\Sexpr{regionStrategyCoverageCI("all", "aap", 0, level)}) to \Sexpr{bestCaseCoverage(1)} (\Sexpr{bestCaseCoverageCI(1, level)}) statewide.

Biweekly rounding yielded regimens similarly non-inferior to the AAP recommendation, with an increase in preventable fraction to \Sexpr{bestCaseCoverage(2)} (\Sexpr{bestCaseCoverageCI(2, level)}) statewide.

However, with monthly rounding, the gains almost completely vanished, with preventable fraction not exceeding \Sexpr{bestCaseCoverage(4)} (\Sexpr{bestCaseCoverageCI(4, level)}). Furthermore, all our regimens with monthly rounding were non-superior to the AAP-recommended regimen in all regions.

\begin{figure}
\begin{center}
\input{\figuresDir/coverageByRoundingRegimen.tex}
\caption[The effect of calendrical rounding of start and end of spatially adjusted RSV prophylaxis regimens on their performance]{The effect of calendrical rounding of start and end of spatially adjusted RSV prophylaxis regimens on their performance, showing that longer rounding intervals diminish the benefits of alternative regimens over the AAP-recommended regimen. Preventable fraction: the fraction of all RSV hospitalizations occurring while the prophylaxis regimen offers protection. Hollow plot markers ($\Circle$, $\square$, and $\triangle$) denote a prophylaxis strategy which, as a result of rounding, is identical to another, and therefore does not provide additional insights.}
\label{fig:coverageByRoundingRegimen}
\end{center}
\end{figure}

%\columnbreak

\paragraph{RSV season onset is slowly moving earlier}

<<onsetOffsetByYear, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/onsetOffsetByYear.tex", figuresDir), width=inlinePlotWidth, height=inlinePlotHeight, pointsize=10)
    data = stateThresholdsByWindow %>% 
        melt(id=c("epiyear")) %>%
        cSplit(c("variable"), c(".")) %>%
        rename(variable=variable_1, pctile=variable_2) %>%
        mutate(variable=factor(variable, levels=c("offset", "onset"))) %>%
        cast(epiyear + variable ~ pctile)

    onsetOffsetAdjust = data.frame(variable=c("onset", "offset"), nudge=c(0, -1))

    limits = data %>% 
        select(variable, median) %>%
        group_by(variable) %>%
        summarize(mid=round(mean(range(median)))) %>%
        inner_join(onsetOffsetAdjust, by="variable") %>%
        mutate(mid=mid + nudge) %>%
        mutate(min=mid - onsetOffsetRange / 2, max=mid + onsetOffsetRange / 2)

    monthBarX = 1998.5 + (1998.5 - 2013.5) * 0.04

    plot = ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=epiyear), width=.3) + 
        geom_point(aes(y=median, x=epiyear), size=.75) +
        geom_blank(data=limits, aes(y=min)) + 
        geom_blank(data=limits, aes(y=max)) + 
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        scale_x_continuous(breaks=seq(1990, 2020, by=5)) +
        scale_y_continuous(minor_breaks=seq(1, 52), breaks=epiWeekBreaks, labels=rep("\\qquad\\qquad", length(epiWeekBreaks)), sec.axis=sec_axis(trans=~., breaks=seq(1, 52))) + 
        annotation_custom(monthGrob(4), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[4], ymax=monthBoundaries$mid[4]) + 
        annotation_custom(monthGrob(5), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[5], ymax=monthBoundaries$mid[5]) + 
        annotation_custom(monthGrob(9), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[9], ymax=monthBoundaries$mid[9]) + 
        annotation_custom(monthGrob(10), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[10], ymax=monthBoundaries$mid[10]) + 
        labs(y=NULL) + 
        theme(
            panel.grid.major.y=element_blank(),
            panel.grid.minor.y=element_line(size=rel(2)),
            axis.title.x=element_blank()
        )

	plotGrob = ggplotGrob(plot)
	plotGrob$layout$clip[grepl("panel", plotGrob$layout$name)] <- "off"
	grid.draw(plotGrob)

    dev.off()
@

Our linear regression of median RSV season onset and offset over three-year periods indicates that season onset has slowly been drifting earlier (Figure \ref{fig:onsetOffsetByYear}), at a rate of \Sexpr{formatDrift(onsetByYear)} (\Sexpr{formatDriftCI(onsetByYear, level)}). It is unclear whether this is a part of a larger pattern, such as a cycle spanning decades.

Season offset, on the other hand, has not been drifting earlier or later (\Sexpr{formatDriftCI(offsetByYear, level)}).

\begin{figure}
\begin{center}
\input{\figuresDir/onsetOffsetByYear.tex}
\caption{Annual variability in RSV season onset and offset in Connecticut}
\label{fig:onsetOffsetByYear}
\end{center}
\end{figure}

\paragraph{Temporally-adjusted prophylaxis regimens are non-superior to others}

<<coverageByRegionWithSliding, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/coverageByRegionRegimenWithSliding.tex", figuresDir), width=pagePlotWidth, height=pagePlotHeight * 1.5, pointsize=10)
    data = unprotectedByCounty %>% 
        filter(county %in% countiesForPlots, rounding==0, !grepl("county", strat))

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        scale_y_continuous(labels=latexPercent, limits=c(0.9, 1), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        scale_color_manual(name="Prophylaxis regimen", labels=legendLabels(stratLabels[c(seq(1, 3), seq(7, 10))]), values=c(rep(c(darkBrown, darkGreen), each=3), darkPurple)) + 
        scale_shape_manual(name="Prophylaxis regimen", labels=legendLabels(stratLabels[c(seq(1, 3), seq(7, 10))]), values=c(16, 15, 17, 16, 15, 17, 16)) + 
        geom_errorbar(
            aes(ymin=frac.lower, ymax=frac.upper, x=county, color=strat, group=strat), 
            width=.3, 
            position=position_dodge(.5)
        ) + 
        geom_point(aes(y=frac.median, x=county, color=strat, shape=strat), position=position_dodge(.5), size=.75) +
        labs(y="Preventable fraction") +
        guides(color=guide_legend(keywidth=3, keyheight=1, ncol=3)) + 
        scale_x_discrete(labels=countyLabels) + 
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            legend.position="bottom"
        )
    dev.off()
@

Compared to our regional all-years analysis, the three regimens based on statewide recent-years analysis produced non-superior results everywhere, and inferior results in the low-RSV counties (Figure \ref{fig:coverageByRegionRegimenWithSliding}).

Due to the non-superiority of regimens based on statewide recent-years analysis, we did not evaluate regimens based on regional recent-years data.

\section{Discussion}

\begin{figure}
\begin{center}
\input{\figuresDir/coverageByRegionRegimenWithSliding.tex}
\caption[Comparison of prophylaxis regimens based on all-years data to those based on recent-years data]{Comparison of prophylaxis regimens based on all-years data to those based on recent-years data, showing temporal analysis adds no benefit over spatial analysis alone. Preventable fraction: fraction of RSV hospitalizations occurring while the prophylaxis regimen offers protection}
\label{fig:coverageByRegionRegimenWithSliding}
\end{center}
\end{figure}

\subsection{Conclusions}

Our analysis confirms the existence of regional variation of RSV season timing in Connecticut (defined by RSV season onset and offset) observable at the county level.

We have also shown a misalignment between the prophylaxis schedule recommended by the AAP and the timing of the RSV season throughout Connecticut, with most counties' RSV season starting before protection from the first dose of prophylaxis takes hold, and ending before protection from the last dose wanes. This misalignment is particularly notable in Fairfield County, due to its early season onset.

Our effort to optimize timing of the prophylaxis regimen (without changing its duration, and therefore its pharmaceutical cost) shows the possibility of a theoretical 1-2\% statewide decrease in RSV-associated hospitalizations among high-risk infants each year, primarily in counties with higher RSV burden (Fairfield, Hartford, and New Haven counties).

The magnitude of that improvement wanes upon encountering practicalities of clinical practice guideline implementation. Nonetheless, with clinical practice guidelines localized at the county level, and the timing of prophylaxis adjusted in each county by less than a whole month from the current AAP guidelines, we still found the potential for $\sim$1\% statewide decrease in RSV-associated hospitalizations among high risk infants.

The prophylaxis regimen with which we were able to attain this improvement is the one in which prophylaxis in each county is timed relative to the midpoint of the RSV season (rounded to two weeks) in that county, which simultaneously reduces unprotected cases early in the season and ineffective prophylaxis late in the season.

Increasing complexity of practice guidelines carries with it increased cost of implementation, such as training of clinicians and other healthcare workers. It is therefore not apparent from our research alone that the potential benefits would outweigh the costs.

Although we found a weak statewide trend in RSV season onset becoming earlier over time, further analysis did not yield any improvements to prophylaxis based on that trend. Since temporal variation in clinical guidelines would lead to a higher implementation burden than spatial variation, we see no reason to base RSV immunoprophylaxis clinical practice guidelines on temporal variation in RSV season timing.

\subsection{Limitations}

Although we were able to reach some conclusions about the impact of calendrical rounding on spatially adjusted prophylaxis regimens, ultimately our analysis is hamstrung by weekly aggregation of hospital admission data. To properly propose and evaluate any fine-tuning to RSV prophylaxis guidelines, we recommend use of daily data.

Throughout, we assumed that RSV-associated hospitalizations are an unbiased estimate of RSV incidence in the general population; in doing so, we implicitly assumed that infectiousness and virulence of RSV are constant over time, for which we have no verification.

Our model of palivizumab immunoprophylaxis assumes a protection period with a hard start at administration and a hard stop at 24 weeks post-administration; the physiologic response to monoclonal antibody immunoprophylaxis is actually tapered on both ends.

Similarly, we assumed the protection by palivizumab is constant at 100\% throughout the protection period; however, not only is breakthrough RSV illness known to occur in patients on palivizumab, but it is an indication for discontinuing the prophylaxis regimen.

In areas of Connecticut associated with a relatively high amount of interstate travel (such as urban centers near the state lines), our analysis --- by only considering residents of Connecticut --- likely underestimates RSV burden.

Finally, towns that see substantial inter-county travel (for example, employment hubs near county lines) likely bias our regional analysis.

\phantomsection
\addcontentsline{toc}{section}{Bibliography}
\printbibliography[title={Bibliography}]

\cleardoublepage
\begin{appendices}

\section{Interval estimation of complex outcomes using penalized basis splines}
\begin{refsection}

\label{supplement:splines}

<<supplement_libraries, cache=FALSE>>=
@

<<supplement, cache=TRUE>>=
@

Penalized basis splines (``$P$-splines'') provide a smoothing mechanism with several qualities desirable in analysis of experimental data, including computational compactness and efficiency, and well-understood statistical behavior. Most importantly for our RSV analysis, the parameters of a $P$-spline approximation have known statistical distributions, which allows interval estimation of arbitrary outcome measures.\cite{Eilers:1996kz}

A simple application of this is our estimation of RSV season onset. Consider, for example, RSV-associated hospitalizations in Tolland county, Connecticut in Figure \ref{fig:tollandCases}. (This is a low-incidence county, which makes it a poor candidate for our main analysis, but its large confidence intervals make it easier to visualize our methods.)

<<tollandCases, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandCases.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    ggplot(tollandObs) +
        theme_light(base_size=plotTextBaseSize) +
        geom_point(aes(x=time, y=rsv.cum.frac), size=.5) +
        scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
        theme(
            panel.grid.major.x=element_blank(),
            legend.title=element_blank(),
            axis.title.y=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
            legend.position="bottom"
        )

    dev.off()
@

\begin{figure}
\begin{center}
\input{\figuresDir/tollandCases.tex}
\caption{Cumulative relative incidence of RSV-associated hospitalizations in Tolland county, Connecticut, aggregated across years \Sexpr{formatYear(startDate)}-\Sexpr{formatYear(endDate)}}
\label{fig:tollandCases}
\end{center}
\end{figure}

<<tollandPredMini, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandPredMini.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    ggplot(tollandPredMini) +
      theme_light(base_size=plotTextBaseSize) +
      geom_line(aes(x=time, y=rsv.cum.frac, group=sim), color="grey") +
      geom_segment(aes(x=-Inf, y=seasonThreshold, xend=+Inf, yend=seasonThreshold), linetype="11", data=data.frame(), size=.375) + 
      geom_point(data=tollandObs, aes(x=time, y=rsv.cum.frac), size=0.5) +
      scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      )

    dev.off()
@

<<tollandOnsetMini, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandOnsetMini.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    ggplot(tollandPredMini) +
      theme_light(base_size=plotTextBaseSize) +
      geom_line(aes(x=time, y=rsv.cum.frac, group=sim), color="grey") +
      geom_segment(aes(x=-Inf, y=seasonThreshold, xend=+Inf, yend=seasonThreshold), linetype="11", data=data.frame(), size=.375) + 
      geom_point(data=tollandObs, aes(x=time, y=rsv.cum.frac), size=0.5) +
      geom_point(data=tollandOnsetMini, aes(x=onset, y=seasonThreshold), size=0.75, shape=17) + 
      geom_segment(data=tollandOnsetMini, aes(x=onset, xend=onset, y=seasonThreshold, yend=0), linetype="11") + 
      scale_y_continuous(limits=c(0, 2*seasonThreshold)) + 
      scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=c(zoomedStartWeek, zoomedEndWeek), expand=c(0, 0)) + 
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      )

    dev.off()
@

\begin{figure}
\begin{center}
\begin{subfigure}{.4\textwidth}
    \begin{center}
    \input{\figuresDir/tollandPredMini}
    \caption{Five randomly sampled $P$-splines approximating the RSV time series.}
    \label{fig:tollandPredMini}
    \end{center}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
    \begin{center}
    \input{\figuresDir/tollandOnsetMini}
    \caption{Season onsets derived from five randomly sampled approximating splines.}
    \label{fig:tollandOnsetMini}
    \end{center}
\end{subfigure}
\end{center}
\caption{Cumulative relative incidence of RSV-associated hospitalizations in Tolland county, Connecticut, with five randomly sampled approximating $P$-splines.}
\end{figure}

We begin the process by defining season onset to occur when cumulative incidence rises above \Sexpr{seasonThreshold * 100}\% of total cases, and obtaining a $P$-spline approximation to our data (using the \texttt{mgcv} R package). This approximation is characterized by a matrix of coefficients, each of which is normally distributed with a known mean and variance.

We randomly sample each coefficient from its distribution, thereby obtaining a specific spline that smoothly approximates our data. Five such randomly sampled splines are shown in Figure \ref{fig:tollandPredMini}. We then calculate RSV season onset for each sampled spline (Figure \ref{fig:tollandOnsetMini}).

Repeating this process many times (\Sexpr{simulations}, in our main analysis) then allows us to estimate the distribution of RSV season onset and therefore make interval estimates. In the case of Tolland county, the result of such season onset sampling is shown in Figure \ref{fig:tollandOnsetFull}.

<<tollandOnsetFull, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandOnsetFull.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    data = tollandPredFull
    splineData = data %>% filter(sim < tollandDisplayNsim)

    ggplot(data) +
      theme_light(base_size=plotTextBaseSize) +
      geom_line(data=splineData, aes(x=time, y=rsv.cum.frac, group=sim), color="gray") +
      geom_segment(aes(x=-Inf, y=seasonThreshold, xend=+Inf, yend=seasonThreshold), linetype="11", data=data.frame(), size=.375) + 
      geom_point(data=tollandObs, aes(x=time, y=rsv.cum.frac), size=0.5) +
      geom_violinh(
        data=tollandOnsetFull,
        aes(x=onset, y=seasonThreshold),
        width=.01,
        fill="gray50", color="black",
        trim=FALSE, draw_quantiles=c(0.025, 0.5, 0.975)
      ) +
      scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, expand=c(0, 0)) +
      scale_y_continuous() +
      coord_cartesian(xlim=c(zoomedStartWeek, zoomedEndWeek), ylim=c(0, 4*seasonThreshold)) +
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      )

      dev.off()
@

\begin{figure}
\begin{center}
\input{\figuresDir/tollandOnsetFull}
\caption{Distribution of approximating $P$-splines ($N=\Sexpr{tollandDisplayNsim}$) and season onset ($N=\Sexpr{tollandNsimFull}$), for cumulative relative incidence of RSV-associated hospitalizations in Tolland county, Connecticut.}
\label{fig:tollandOnsetFull}
\end{center}
\end{figure}

We use the same process to estimate RSV season offset; the process can be generalized to any scalar outcome measure of an infectious disease outbreak or season.

Furthermore, because our approach involves sampling approximations to the observed time series as a whole, rather than estimating the time series at isolated points in time, we are able to produce interval estimates of arbitrarily complex outcome measures.

We take advantage of this in our estimation of the preventable fraction of RSV cases. Here we begin by defining a preventable case of RSV to be a case of RSV that occurs while prophylaxis offers protection, and we define the preventable fraction of a prophylaxis administration strategy to be the ratio of preventable cases under that prophylaxis strategy to the total cases of RSV.

As before, we then sample $P$-spline coefficients to obtain a specific $P$-spline approximation, and compute preventable fraction for each sampled $P$-spline in order to obtain the distribution of preventable fractions.

<<tollandPredSingle, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandPredSingle.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    data = tollandPredSingle

    thresholds = data %>%
    	mutate(ppx=strat(time)) %>%
    	filter(ppx > 0) %>%
    	do((function(df) {
    	  df = df %>% arrange(time)
    		data.frame(
    			ppx.start=min(df$time),
    			ppx.end=max(df$time),
    			unprotected.start=first(df$rsv.cum.frac),
    			unprotected.end=last(df$rsv.cum.frac)
    		)
    	})(.)) %>%
    	ungroup() %>%
    	as.data.frame()

    ggplot(data) +
      theme_light(base_size=plotTextBaseSize) +
      geom_line(aes(x=time, y=rsv.cum.frac, group=sim), color="grey") +
      geom_segment(data=thresholds, aes(x=ppx.start, y=unprotected.start, xend=ppx.start, yend=+Inf), linetype="11", size=.375) + 
      geom_segment(data=thresholds, aes(x=ppx.end, y=unprotected.end, xend=ppx.end, yend=+Inf), linetype="11", size=.375) + 
      geom_segment(data=thresholds, aes(x=ppx.start, y=unprotected.start, xend=+Inf, yend=unprotected.start), linetype="11", size=.375) + 
      geom_segment(data=thresholds, aes(x=ppx.end, y=unprotected.end, xend=+Inf, yend=unprotected.end), linetype="11", size=.375) + 
      geom_point(data=tollandObs, aes(x=time, y=rsv.cum.frac, size=0.5), size=0.5) +
    	geom_rect(data=thresholds, aes(xmin=ppx.start, xmax=ppx.end, ymin=1, ymax=1.1), size=.375) + 
    	geom_text(data=thresholds, aes(x=mean(c(ppx.start, ppx.end)), y=1.05, label="Prophylaxis window"), color="white", size=1.75) + 
    	geom_rect(data=thresholds, aes(xmin=max(monthBoundaries$max) - 3, xmax=max(monthBoundaries$max), ymin=unprotected.start, ymax=unprotected.end), size=.375) + 
    	geom_text(data=thresholds, aes(x=max(monthBoundaries$max) - 1.5, y=mean(c(unprotected.start, unprotected.end)), label="$\\Delta$ = preventable fraction", angle=90, lineheight=0.5), color="white", size=1.75) + 
      scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      )

    dev.off()
@

<<tollandPreventableFractionFull, results='hide', warning=FALSE>>=
    tikz(sprintf("%s/tollandPreventableFractionFull.tex", figuresDir), width=pagePlotWidth * 0.4, height=pagePlotHeight, pointsize=10)

    data = tollandPredFull
    splineData = data %>% filter(sim < tollandDisplayNsim)

    splinePlot = ggplot(data) +
      theme_light(base_size=plotTextBaseSize) +
      geom_line(data=splineData, aes(x=time, y=rsv.cum.frac, group=sim), color="gray") +
      geom_segment(data=thresholds, aes(x=ppx.start, y=-Inf, xend=ppx.start, yend=+Inf), linetype="11", size=.375) + 
      geom_segment(data=thresholds, aes(x=ppx.end, y=-Inf, xend=ppx.end, yend=+Inf), linetype="11", size=.375) + 
    	geom_rect(data=thresholds, aes(xmin=ppx.start, xmax=ppx.end, ymin=1, ymax=1.1), size=.375) + 
    	geom_text(data=thresholds, aes(x=mean(c(ppx.start, ppx.end)), y=1.05, label="Prophylaxis window"), color="white", size=1.75) + 
      geom_point(data=tollandObs, aes(x=time, y=rsv.cum.frac), size=0.5) +
      scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, expand=c(0, 0)) +
      scale_y_continuous() +
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      )

    violinPlot <- ggplot(tollandFractionFull) +
      theme_light(base_size=plotTextBaseSize) +
      geom_violin(
        aes(x=0, y=unprotected),
        width=.01,
        fill="gray50", color="black",
        trim=FALSE, draw_quantiles=c(0.025, 0.5, 0.975)
      ) +
      scale_x_continuous() +
      scale_y_continuous(limits=c(0, 1)) +
      theme(
        panel.grid.major.x=element_blank(),
        legend.title=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
        legend.position="bottom"
      ) 
    grid.arrange(splinePlot, violinPlot, layout_matrix=rbind(c(rep(1, 4), 2)))

    dev.off()
@

\begin{figure}
\begin{center}
\begin{subfigure}{.4\textwidth}
    \begin{center}
    \input{\figuresDir/tollandPredSingle}
    \caption{Preventable fraction derived from a randomly sampled $P$-spline approximating the RSV time series.}
    \label{fig:tollandPredMini}
    \end{center}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
    \begin{center}
    \input{\figuresDir/tollandPreventableFractionFull}
    \caption{Distribution of approximating $P$-splines ($N=\Sexpr{tollandDisplayNsim}$) and preventable fractions ($N=\Sexpr{tollandNsimFull}$) of RSV under the AAP prophylaxis guidelines  in Tolland county, Connecticut.}
    \label{fig:tollandOnsetMini}
    \end{center}
\end{subfigure}
\end{center}
\caption{Cumulative relative incidence of RSV-associated hospitalizations in Tolland county, Connecticut, with five randomly sampled approximating $P$-splines.}
\end{figure}

We developed a custom R package for estimation of arbitrary characteristics of infectious disease outbreaks or seasons (as long as they are amenable to generalized additive modeling); the package is available at \href{https://github.com/airbornemint/outbreak-inference}{\nolinkurl{github.com/airbornemint/outbreak-inference}}.

\printbibliography[heading=subbibliography]
\end{refsection}

\end{appendices}

\end{document}