### 2020-10-04 paper/structure.tex
Refactored for SciRep

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\usepackage[english]{babel} % English language hyphenation

\usepackage{microtype} % Better typography

\usepackage{amsmath,amsfonts,amsthm} % Math packages for equations

\usepackage[labelfont=bf, textfont=it]{caption} % Custom captions under/above tables and figures
\usepackage{subcaption}
\DeclareCaptionType{supplementalfigure}[Supplemental Figure][List of supplemental figures]

\usepackage{booktabs} % Horizontal rules in tables

\usepackage{graphicx} % Required for adding images

\usepackage{enumitem} % Required for customising lists
\setlist{noitemsep} % Remove spacing between bullet/numbered list elements

\usepackage{sectsty} % Enables custom section titles
\allsectionsfont{\usefont{OT1}{phv}{b}{n}} % Change the font of all section commands (Helvetica)